### 2020-04-30 documents/paper.Rmd
outline the ms

---
title: "The Geometry of Archaeological Thought: ..."
output: bookdown::html_document2
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

# Background

- previous efforts at the history of archaeology (e.g. Bruce Trigger's book)
- previous efforts at text analysis in archaeology (e.g. bibliometrics, citation networks)

# Methods

- source of the data
- preparation steps: data cleaning (Cite the quanteda pkg, and things they cite)
- word embeddings (cite our readings)
- temporal referencing plots (cite our readings)

# Results

- basic details (e.g. word and abstract counts)
- word frequency time series
- TR plots

# Discussion

- ...

# Conclusion

- ...

# Acknowledgements

# References

<!--- use the bib file -->

### 2020-05-06 code/003-basic-word-stats-and-plots.R
make the trends show more clearly on the time series plots

@@ -1,5 +1,3 @@
-
-
 # Words per year
 
 library(tidyverse)
@@ -63,22 +61,27 @@ dfm_keywords_tbl <-
 dfm_keywords_tbl_prop <- 
 dfm_keywords_tbl %>% 
   left_join(all_txts_c_summary) %>% 
-  mutate(prop = n / Tokens )
+  mutate(prop = n / Tokens ) %>% 
+  group_by(keyword) %>% 
+  mutate(sum_the_word = sum(n)) %>% 
+  mutate(keyword_n = str_c(keyword, " (n = ", sum_the_word, ")"))
 
 # plot of keywords as a proportion of all words per year
 ggplot(data = dfm_keywords_tbl_prop, 
        aes(x = year , 
            y = prop)) +
-  geom_col() +
-  facet_wrap( ~ keyword, 
-              ncol = 1) + #,
-             # scales = "free_y") +
+  geom_point(alpha = 0.3) +
+  geom_smooth(span = 0.475,
+              se = FALSE) +
+  facet_wrap( ~ keyword_n, 
+              ncol = 1,
+             scales = "free_y") +
   scale_x_continuous(labels = c(seq(1960, 2020, 2)), 
                      breaks = seq(1960, 2020, 2),
                      name = "Year") +
   scale_y_continuous(name = "Proportion of all words per year",
                      labels = scales::comma) +
-  theme_bw() +
+  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, 
                                    vjust = 0.5)) 
 

### 2020-05-06 code/003-exploring-basic-word-stats-and-plots.R

@@ -29,10 +29,10 @@ all_txts_c_dtm <-
 
 #----- change these words
 keywords <-
-  c('model',
-    "explanation",
-    "theory",
-    "hypothesis",
+  c('agency',
+    "phenomena",
+    "action",
+    "affect",
     "inference",
     "mechanism")
 #----

### 2020-05-06 documents/paper.Rmd

@@
 - source of the data
+We collected abstracts from Society of American Archaeology (SAA) annual meeting programs from 1964 to 2020.
+The total number of abstracts are 68,176 from 40 annual meetings excluding some years without available abstracts.
+The source of abstract includes papers, posters, workshop posters, symposium represented by paragraphs.
+The annual programs are available online at https://www.saa.org/annual-meeting/annual-meeting-archives/program-archives.
+The original format of abstracts on websites includes scanned images for early years and accessible PDF after 2004.
+
 - preparation steps: data cleaning (Cite the quanteda pkg, and things they cite)
+
 - word embeddings (cite our readings)

### 2020-05-06 code/003-basic-word-stats-and-plots.R
number of abstracts for all years

@@ -79,7 +79,7 @@ ggplot(data = dfm_keywords_tbl_prop,
   scale_x_continuous(labels = c(seq(1960, 2020, 2)), 
                      breaks = seq(1960, 2020, 2),
                      name = "Year") +
-  scale_y_continuous(name = "Proportion of all words per year",
+  scale_y_continuous(name = "Proportion of all words in all abstracts per year",
                      labels = scales::comma) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, 
@@ -108,13 +108,30 @@ for(i in seq_len(length(all_txts_text_pdf))){
 }
 
 number_of_abstracts_per_year <- 
-tibble(year = names(all_txts_text_pdf),
+tibble(year = parse_number(names(all_txts_text_pdf)),
        number_of_abstracts = unlist(all_txts_text_pdf_tally))
 
-ggplot(number_of_abstracts_per_year,
+# also get count of abstracts from image scans also
+number_of_abstracts_per_year_all_years <- 
+  readxl::read_excel(here::here("data/raw-data/saa-abstracts-tally.xlsx")) %>% 
+  filter(image_or_text == "image") %>% 
+  select(year, number_of_abstracts) %>% 
+  bind_rows(number_of_abstracts_per_year)
+
+# plot abstracts for all years
+ggplot(number_of_abstracts_per_year_all_years,
        aes(year, 
            number_of_abstracts)) +
-  geom_col()
+  geom_col() +
+  scale_x_continuous(labels = c(seq(1960, 2020, 2)), 
+                     breaks = seq(1960, 2020, 2),
+                     name = "Year") +
+  scale_y_continuous(name = "Number of abstracts per year",
+                     labels = scales::comma) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, 
+                                   vjust = 0.5)) 
+
 
 
 

### 2020-05-06 code/004-excersize-abstract-count.R
add abstract code

@@ -1,0 +1,16 @@
+library(ggplot2)
+library(tidyverse)
+
+abs <- readxl::read_excel(here::here("data", "raw-data", "saa-abstracts-tally.xlsx"))
+
+year_with_abs <-
+  abs %>% 
+  filter(!is.na(`number_of_abstracts`))
+
+ggplot(year_with_abs,
+       aes(x = year,
+           y = `number_of_abstracts`)) +
+  geom_bar(stat="identity") +
+  theme_minimal() +
+  scale_x_continuous(breaks = seq(from =1962, to = 2020, by =2)) +
+  theme(axis.text.x = element_text(angle = 90))

### 2020-05-06 code/003-basic-word-stats-and-plots.R
explore words

@@ -45,7 +45,7 @@ all_txts_c_dtm <-
 # Our key words over time
   
 keywords <- 
-  c('model', "explanation", "theory", "hypothesis", "inference", "mechanism")
+  c("experience", "process", "community", "dynamics", "context", "compare")
 
 dfm_keywords <- 
   dfm_select(all_txts_c_dtm, 
@@ -112,7 +112,4 @@ ggplot(number_of_abstracts_per_year,
        aes(year, 
            number_of_abstracts)) +
   geom_col()
-
-
-
 

### 2020-05-06 code/003-exploring-basic-word-stats-and-plots.R

@@ -29,12 +29,12 @@ all_txts_c_dtm <-
 
 #----- change these words
 keywords <-
-  c('model',
-    "explanation",
-    "theory",
-    "hypothesis",
-    "inference",
-    "mechanism")
+  c("experience", 
+    "process", 
+    "community", 
+    "dynamics", 
+    "context", 
+    "compare")
 #----
 
 dfm_keywords <- 

### 2020-05-07 documents/paper.Rmd
fix 1985 text, which had a problem with PDF -> PNG conversion

@@
 ---
-title: "The Geometry of Archaeological Thought: ..."
+title: "The Geometry of Archaeological Thought: Analysing discplinary change using word embeddings from conference abstracts"
 output: bookdown::html_document2
+bibliography: bibliography.bib
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 # Introduction
+
+The study of disciplinary change traditionally relies on close reading of a large number of publications, manually sorting them into pre-defined categories and making connections between these groups and external influences [e.g.
+@trigger1989history].
+These methods are time-consuming and expensive, which excludes many people from participating.
+They are also vulnerable to biases
+
+computational approaches that could scan the entire corpus to generate topics in a bottom-up fashion, based on the semantic content of the publications themselves.
+These approaches indeed make it possible to analyze more comprehensively and more systematically the contents of large corpora of full-text documents, be it in history and sociology
 

### 2020-05-07 code/001-PDF-page-images-to-txt.R

@@ -11,7 +11,7 @@ abstracts_tally <- readxl::read_excel(here::here("data/raw-data/saa-abstracts-ta
 abstracts_that_are_page_images <- 
 abstracts_tally %>% 
   filter(image_or_text == "image") %>% 
-  filter( year %in% 1993:1994)
+  filter( year %in% 1985) # checking just one year that seems odd, remove this line to do all years
 
 # OCR abstracts that are images
 
@@ -44,13 +44,14 @@ library(magick)
 
 abstracts_that_are_page_images_two_page_spread <- 
   abstracts_that_are_page_images %>% 
-  filter(single_or_double_page_spread == "double")
+  filter(single_or_double_page_spread == "double") %>% 
+  filter( year %in% 1985) # checking just one year that seems odd, remove this line to do all years
 
 abstracts_that_are_page_images_two_page_spread_paths <- 
   paste0(here::here("data/derived-data/abstracts-ocr/"), 
                 abstracts_that_are_page_images_two_page_spread$year)
 
-# only work on those directories that actuall exist currently 
+# only work on those directories that actually exist currently 
 abstracts_that_are_page_images_two_page_spread_paths_existing <- 
   abstracts_that_are_page_images_two_page_spread_paths[dir.exists(abstracts_that_are_page_images_two_page_spread_paths)]
 

### 2020-05-07 code/003-basic-word-stats-and-plots.R

@@ -14,6 +14,16 @@ abstract_txt_files <-
 all_txts <- tolower(readtext::readtext(abstract_txt_files))
 names(all_txts) <- map_chr(names(all_txts), ~str_match_all(.x, "\\d{4}")[[1]][1,1])
 
+# word count checking
+all_txts_word_count_per_year <- 
+map_int(all_txts, 
+       stringi::stri_count_words) %>% 
+  stack %>% 
+  as_tibble() %>% 
+  rename(words = values,
+         year = ind) %>% 
+  mutate(year = as.numeric(as.character(year)))
+
 # count all words for each year
 all_txts_c <- corpus(all_txts)
 all_txts_c_summary <- 
@@ -26,6 +36,7 @@ all_txts_c_dtm <-
       remove_punct = TRUE)
 
 # plot of all words in each year
+plot_all_words <- 
   ggplot(data = all_txts_c_summary, 
          aes(x = year , 
              y = Tokens / 1000, 
@@ -34,9 +45,9 @@ all_txts_c_dtm <-
   scale_x_continuous(labels = c(seq(1940, 2020, 2)), 
                      breaks = seq(1940, 2020, 2),
                      name = "Year") +
-  scale_y_continuous(name = "Total word count (x 1,000)",
+  scale_y_continuous(name = "Total word count per year (x 1,000)",
                      labels = scales::comma) +
-  theme_bw() +
+  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, 
                                    vjust = 0.5)) 
   
@@ -119,6 +130,7 @@ number_of_abstracts_per_year_all_years <-
   bind_rows(number_of_abstracts_per_year)
 
 # plot abstracts for all years
+plot_all_abstracts <- 
 ggplot(number_of_abstracts_per_year_all_years,
        aes(year, 
            number_of_abstracts)) +
@@ -131,6 +143,9 @@ ggplot(number_of_abstracts_per_year_all_years,
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, 
                                    vjust = 0.5)) 
+
+library(patchwork)
+plot_all_words / plot_all_abstracts
 
 
 

### 2020-05-07 documents/paper.Rmd
work on method section and create bib file

--- 2020-05-06 documents/paper.Rmd
@@
 The annual programs are available online at https://www.saa.org/annual-meeting/annual-meeting-archives/program-archives.
-The original format of abstracts on websites includes scanned images for early years and accessible PDF after 2004.
+The original format of abstracts on websites includes non-searchable PDF documents coming from images for early years and searchable PDF documents after 2005.
 
 - preparation steps: data cleaning (Cite the quanteda pkg, and things they cite)
+For image-based pdfs, we convert them into PNG files, clean up the images using Magick package, and then extract the text using Tesseract package [@Kay2007] for optical character recognition (OCR).
+We obtain 95-98 percent accuracy of OCR on a word level across years according to the results of error rate for each year.
+After the text are ready, we use quanteda package for text managing and analyzing [@Benoit2018].
+We clean up the text by removing all punctuation and stop words, such as "the", "an", "a", that might have some impacts on our results.
 
 - word embeddings (cite our readings)
+Word embeddings is useful tool based machine learning framework for measurement, quantifying, and comparing the semantic relations between words, which enables a wide range of application for temporal changes in word relationships related to the sociocultural domain [@Garg2018; @Kozlowski2019].
+There are two major word-level pretrained word embeddings, Weod2Vec and Glove according to different methods [@Rong2014; @Pennington2014].
+Since pretrained word embeddings are trained on large corpus such as Google News, they can capture the semantic meaning  of a word very well.
+
+In this article, we use text2vec package [@Selivanov2016] to train the word embeddings using GloVe model, which calculates the occurrence of a pair of words and creates the co-occurrence matrix.
+
 - temporal referencing plots (cite our readings)

### 2020-05-07 code/003-basic-word-stats-and-plots.R
explore words

@@ -43,7 +43,12 @@ all_txts_c_dtm <-
 # Our key words over time
   
 keywords <- 
-  c("experience", "process", "community", "dynamics", "context", "compare")
+  c("identity", 
+    "power", 
+    "awareness", 
+    "transformation", 
+    "presentation", 
+    "cognition")
 
 dfm_keywords <- 
   dfm_select(all_txts_c_dtm, 

### 2020-05-07 documents/paper.Rmd
correct some typos

@@
 - word embeddings (cite our readings)
-Word embeddings is useful tool based machine learning framework for measurement, quantifying, and comparing the semantic relations between words, which enables a wide range of application for temporal changes in word relationships related to the sociocultural domain [@Garg2018; @Kozlowski2019].
-There are two major word-level pretrained word embeddings, Weod2Vec and Glove according to different methods [@Rong2014; @Pennington2014].
+Word embeddings is useful tool based machine learning framework for measurement, quantifying, and comparing the semantic relations between words, which enables a wide range of application for temporal changes in word relationships related to the sociocultural interests [@Garg2018; @Kozlowski2019].
+There are two major word-level pretrained word embeddings, Word2Vec and Glove according to different methods used for training [@Rong2014; @Pennington2014].
 Since pretrained word embeddings are trained on large corpus such as Google News, they can capture the semantic meaning of a word very well.

### 2020-05-07 documents/paper.Rmd
stash

--- 2020-05-07 documents/paper.Rmd
@@
 @trigger1989history].
-These methods are time-consuming and expensive, which excludes many people from participating.
-They are also vulnerable to biases
+These methods are time-consuming and expensive (because they depend on full access to a large library of texts), which limits participation in this aspect of archaeology.
+Computational approaches to analyzing text that can process many publications quickly to generate topics in a bottom-up fashion offer a new way to investigate disciplinary history.
+This has the advantage of increasing and diversifying participation in study of the history of archaeology, as well as  more comprehensive, systematic, and reproducible analyses of the contents of large number of full-text documents.
+The methods and data we present here offer empirical basis for what might otherwise be informal claims about the discipline and its history.
+They are also the type of data that may prompt discussion about the directions the discipline might — or could — take.
 
-computational approaches that could scan the entire corpus to generate topics in a bottom-up fashion, based on the semantic content of the publications themselves.
-These approaches indeed make it possible to analyze more comprehensively and more systematically the contents of large corpora of full-text documents, be it in history and sociology
+In this paper we present a dataset of all available abstracts of papers and posters presented at annual meetings of the Society of American Archaeology.
+We demonstrate the use of word embeddings, a commonly used computational tool in natural language processing (NLP) and machine learning, as a method to measure, quantify, and compare changes in the way archaeologists write.
+We present the temporal reference plot as a method for visualizing disciplinary change in archaeology.
+As a case study, we use these data and methods to explore how archaeologists' writing about explaining and theorizing the past have changed over time.
 
 # Background
+
+Research into the history of archaeology is often organised into two major phases.
+An early internalist phase, where the story of development of the field is told with limited reference to the influence of political, social, religious, and economic institutions.
+In the 1980s an external, critical history phase began, seeking to understand the interactions between archaeological practice and its political, economic and social contexts.
+
+Among historians of archaeology there are two major phases in  ‘critical history of archaeology’ in the 1980s had been characterized by the replacement of an
 
@@
 The annual programs are available online at https://www.saa.org/annual-meeting/annual-meeting-archives/program-archives.
-The original format of abstracts on websites includes non-searchable PDF documents coming from images for early years and searchable PDF documents after 2005.
+The original format of abstracts on websites includes scanned images for early years and accessible PDF after 2004.
 
 - preparation steps: data cleaning (Cite the quanteda pkg, and things they cite)
-For image-based pdfs, we convert them into PNG files, clean up the images using Magick package, and then extract the text using Tesseract package [@Kay2007] for optical character recognition (OCR).
-We obtain 95-98 percent accuracy of OCR on a word level across years according to the results of error rate for each year.
-After the text are ready, we use quanteda package for text managing and analyzing [@Benoit2018].
-We clean up the text by removing all punctuation and stop words, such as "the", "an", "a", that might have some impacts on our results.
 
 - word embeddings (cite our readings)
-Word embeddings is useful tool based machine learning framework for measurement, quantifying, and comparing the semantic relations between words, which enables a wide range of application for temporal changes in word relationships related to the sociocultural domain [@Garg2018; @Kozlowski2019].
-There are two major word-level pretrained word embeddings, Weod2Vec and Glove according to different methods [@Rong2014; @Pennington2014].
-Since pretrained word embeddings are trained on large corpus such as Google News, they can capture the semantic meaning  of a word very well.
-
-In this article, we use text2vec package [@Selivanov2016] to train the word embeddings using GloVe model, which calculates the occurrence of a pair of words and creates the co-occurrence matrix.
-
 - temporal referencing plots (cite our readings)

### 2020-05-07 code/001-PDF-page-images-to-txt.R

@@ -11,7 +11,7 @@ abstracts_tally <- readxl::read_excel(here::here("data/raw-data/saa-abstracts-ta
 abstracts_that_are_page_images <- 
 abstracts_tally %>% 
   filter(image_or_text == "image") %>% 
-  filter( year %in% 1985) # checking just one year that seems odd, remove this line to do all years
+  filter( year %in% 1982) # checking just one year that seems odd, remove this line to do all years
 
 # OCR abstracts that are images
 
@@ -45,7 +45,7 @@ library(magick)
 abstracts_that_are_page_images_two_page_spread <- 
   abstracts_that_are_page_images %>% 
   filter(single_or_double_page_spread == "double") %>% 
-  filter( year %in% 1985) # checking just one year that seems odd, remove this line to do all years
+  filter( year %in% 1982) # checking just one year that seems odd, remove this line to do all years
 
 abstracts_that_are_page_images_two_page_spread_paths <- 
   paste0(here::here("data/derived-data/abstracts-ocr/"), 

### 2020-05-07 code/004-excersize-abstract-count.R

@@ -1,16 +1,0 @@
-library(ggplot2)
-library(tidyverse)
-
-abs <- readxl::read_excel(here::here("data", "raw-data", "saa-abstracts-tally.xlsx"))
-
-year_with_abs <-
-  abs %>% 
-  filter(!is.na(`number_of_abstracts`))
-
-ggplot(year_with_abs,
-       aes(x = year,
-           y = `number_of_abstracts`)) +
-  geom_bar(stat="identity") +
-  theme_minimal() +
-  scale_x_continuous(breaks = seq(from =1962, to = 2020, by =2)) +
-  theme(axis.text.x = element_text(angle = 90))

### 2020-05-07 documents/paper.Rmd

@@
 The annual programs are available online at https://www.saa.org/annual-meeting/annual-meeting-archives/program-archives.
-The original format of abstracts on websites includes scanned images for early years and accessible PDF after 2004.
+The original format of abstracts on websites includes non-searchable PDF documents coming from images for early years and searchable PDF documents after 2005.
 
 - preparation steps: data cleaning (Cite the quanteda pkg, and things they cite)
+For image-based pdfs, we convert them into PNG files, clean up the images using Magick package, and then extract the text using Tesseract package [@Kay2007] for optical character recognition (OCR).
+We obtain 95-98 percent accuracy of OCR on a word level across years according to the results of error rate for each year.
+After the text are ready, we use quanteda package for text managing and analyzing [@Benoit2018].
+We clean up the text by removing all punctuation and stop words, such as "the", "an", "a", that might have some impacts on our results.
 
 - word embeddings (cite our readings)
+Word embeddings is useful tool based machine learning framework for measurement, quantifying, and comparing the semantic relations between words, which enables a wide range of application for temporal changes in word relationships related to the sociocultural interests [@Garg2018; @Kozlowski2019].
+There are two major word-level pretrained word embeddings, Word2Vec and Glove according to different methods used for training [@Rong2014; @Pennington2014].
+Since pretrained word embeddings are trained on large corpus such as Google News, they can capture the semantic meaning of a word very well.
+
+In this article, we use text2vec package [@Selivanov2016] to train the word embeddings using GloVe model, which calculates the occurrence of a pair of words and creates the co-occurrence matrix.
+
 - temporal referencing plots (cite our readings)

### 2020-05-08 documents/paper.Rmd
exporing word prop changes by two periods

--- 2020-05-07 documents/paper.Rmd
@@
 output: bookdown::html_document2
-bibliography: bibliography.bib
+bibliography: references.bib
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 In this paper we present a dataset of all available abstracts of papers and posters presented at annual meetings of the Society of American Archaeology.
-We demonstrate the use of word embeddings, a commonly used computational tool in natural language processing (NLP) and machine learning, as a method to measure, quantify, and compare changes in the way archaeologists write.
+We demonstrate the use of word embeddings, a commonly used computational tool in natural language processing (NLP) and machine learning, as a method to measure, quantify, and compare changes in the way archaeologists write cf.
+[@Garg2018; @Kozlowski2019].
 We present the temporal reference plot as a method for visualizing disciplinary change in archaeology.
@@
 - source of the data
-We collected abstracts from Society of American Archaeology (SAA) annual meeting programs from 1964 to 2020.
-The total number of abstracts are 68,176 from 40 annual meetings excluding some years without available abstracts.
-The source of abstract includes papers, posters, workshop posters, symposium represented by paragraphs.
-The annual programs are available online at https://www.saa.org/annual-meeting/annual-meeting-archives/program-archives.
-The original format of abstracts on websites includes scanned images for early years and accessible PDF after 2004.
 
-- preparation steps: data cleaning (Cite the quanteda pkg, and things they cite)
+The data for our study comes from the text of oral and poster presentation abstracts presented at the annual meetings of the Society of American Archaeology (SAA).
+The earliest meeting in our data is 19XX, and we have some years missing, especially XXXX-XXXX.
+The total number of abstracts is 68,176 (xxx words) from 40 annual meetings.
+We obtained PDF files of annual programs from https://www.saa.org/annual-meeting/annual-meeting-archives/program-archives.
+
+The PDFs for the years XXXX-XXXX contain images of pages, and after 2005 the PDFs are searchable text.
+We converted the page images from PDF to PNG files using ImageMagick software, and then used optical character recognition to extract text from the PNG files with the Tesseract package [@Kay2007].
+We obtained an acceptable rate of 95-98% word-level OCR accuracy across the PDFs [citation].
+After OCR, we used the quanteda package  [@Benoit2018] to remove punctuation, numbers, and stopwords, such as "the", "an", "a", [specifically why do we remove stopwords?].
+
+- something about basic steps
+-- transform text into dfm, define dfm...
 
 - word embeddings (cite our readings)
+-- a few sentences to introduce the reader to how WE have been used in other recent studies.
+this is a little bit background.
+'...recently used in many social science fields to...' tell the reader what was the finding of some key studies using WE that have inspired us.
+
+<!-- There are two major word-level pretrained word embeddings, Word2Vec and Glove according to different methods used for training [@Rong2014; @Pennington2014]. -->
+
+-- then some more detail about the mechanics of the method, as we are using it. Say how what we did works in general.
+
+In this article, we trained word embeddings with the GloVe model using th text2vec package [@Selivanov2016]. The GloVe model calculates ... creates the co-occurrence matrix.
+
 - temporal referencing plots (cite our readings)

### 2020-05-08 code/0041-basic-word-stats-and-plots.R

@@ -1,0 +1,135 @@
+df <- all_txts_c_dtm
+
+dont_care <- 
+  c('university', 'abstracts', 'meeting', 
+    'annual', 'paper', 'chair', 'across', 
+    'presented', 'session', 'college', 'area',
+    'discussed', 'this', 'that', 'there', 'which',
+    'their', 'poster', 'using', 'through', 'into',
+    'some', 'from', 'been', 'discuss', 'while',
+    'over')
+
+df_tbl <- 
+  convert(df, to = "data.frame") %>% 
+  pivot_longer(-document, 
+               names_to = "word",
+               values_to = "n") %>% 
+  mutate(year = parse_number(document)) %>% 
+  filter(word %in% names(data_int_syllables)) %>% 
+  filter(str_length(word) > 3) %>% 
+  filter(!word %in% dont_care)
+  
+# compute proportion of all words per year
+df_tbl_prop <- 
+  df_tbl %>% 
+  left_join(all_txts_c_summary) %>% 
+  mutate(prop = n / Tokens ) %>% 
+  filter(prop > 0) %>% 
+  group_by(word) %>% 
+  mutate(sum_the_word = sum(n)) %>% 
+  mutate(keyword_n = str_c(word, " (n = ", sum_the_word, ")"))
+
+# want to find words with the greatest difference in 
+# proportion between two time periods 
+
+df_tbl_prop_two_groups <- 
+  df_tbl_prop %>% 
+  mutate(year_group = ifelse(document < 2004, "early", "late"))
+
+df_tbl_prop_two_groups_diff <- 
+df_tbl_prop_two_groups %>% 
+  group_by(word, year_group) %>% 
+  summarise(mean_prop = mean(prop, na.rm = TRUE)) %>% 
+  pivot_wider(names_from = year_group,
+              values_from = mean_prop) %>% 
+  drop_na() %>% 
+  mutate(diff = late - early)
+
+df_tbl_prop_two_groups_diff_interesting <- 
+df_tbl_prop_two_groups_diff %>% 
+  filter(diff >= 0.00001 | diff <= -0.00001)
+
+df_tbl_prop_two_groups_diff_interesting_top <- 
+  df_tbl_prop_two_groups_diff_interesting %>% 
+  dplyr::arrange(desc(diff)) %>% 
+  head(30)
+
+df_tbl_prop_two_groups_diff_interesting_bottom <- 
[…]

### 2020-05-08 code/003-exploring-basic-word-stats-and-plots.R
commit changes in word exploration before merging

@@ -29,7 +29,7 @@ all_txts_c_dtm <-
 
 #----- change these words
 keywords <-
-  c('agency',
+  c("agency",
     "phenomena",
     "action",
     "affect",
@@ -61,7 +61,7 @@ ggplot(data = dfm_keywords_tbl_prop,
            y = prop)) +
   geom_col() +
   facet_wrap( ~ keyword, 
-              ncol = 1,
+              ncol = 1, 
              scales = "free_y") +
   scale_x_continuous(labels = c(seq(1960, 2020, 2)), 
                      breaks = seq(1960, 2020, 2),

### 2020-05-08 documents/paper.Rmd
add code chunk for in-line code and one references

@@
 
+```{r read-data}
+saa_data <- readxl::read_excel(here::here("data", "raw-data", "saa-abstracts-tally.xlsx"))
+abstract_num <- sum(saa_data$number_of_abstracts, na.rm = TRUE)
+year_num <- length(which(!is.na(saa_data$number_of_abstracts)))
+```
+
 The data for our study comes from the text of oral and poster presentation abstracts presented at the annual meetings of the Society of American Archaeology (SAA).
-The earliest meeting in our data is 19XX, and we have some years missing, especially XXXX-XXXX.
-The total number of abstracts is 68,176 (xxx words) from 40 annual meetings.
+The earliest meeting in our data is 1962, and we have some years missing, especially 1995-2003.
+The total number of abstracts is `r abstract_num` (xxx words) from `r year_num` annual meetings.
 We obtained PDF files of annual programs from https://www.saa.org/annual-meeting/annual-meeting-archives/program-archives.
 
-The PDFs for the years XXXX-XXXX contain images of pages, and after 2005 the PDFs are searchable text.
+The PDFs for the years 1962-1994 contain images of pages, and after 2005 the PDFs are searchable text.
 We converted the page images from PDF to PNG files using ImageMagick software, and then used optical character recognition to extract text from the PNG files with the Tesseract package [@Kay2007].
-We obtained an acceptable rate of 95-98% word-level OCR accuracy across the PDFs [citation].
-After OCR, we used the quanteda package  [@Benoit2018] to remove punctuation, numbers, and stopwords, such as "the", "an", "a", [specifically why do we remove stopwords?].
+We obtained an acceptable rate of 95-98% word-level OCR accuracy across the PDFs [@Holley2009].
+After OCR, we used the quanteda package  [@Benoit2018] to remove punctuation, numbers, and stopwords, such as "the", "an", "a", with little or no semantic meanings.
 

### 2020-05-10 documents/paper.Rmd
editing

@@
 The data for our study comes from the text of oral and poster presentation abstracts presented at the annual meetings of the Society of American Archaeology (SAA).
-The earliest meeting in our data is 1962, and we have some years missing, especially 1995-2003.
-The total number of abstracts is `r abstract_num` (xxx words) from `r year_num` annual meetings.
+The earliest meeting in our data is 1962, and we have some years missing, especially 1995-2003 where PDF files were not available to us.
+The total number of abstracts is `r prettyNum(abstract_num, big.mark = ",")` (xxx words) from `r year_num` annual meetings.
 We obtained PDF files of annual programs from https://www.saa.org/annual-meeting/annual-meeting-archives/program-archives.

### 2020-05-14 documents/paper.Rmd
cminor edit on method

--- 2020-05-08 documents/paper.Rmd
@@
 We obtained an acceptable rate of 95-98% word-level OCR accuracy across the PDFs [@Holley2009].
-After OCR, we used the quanteda package  [@Benoit2018] to remove punctuation, numbers, and stopwords, such as "the", "an", "a", with little or no semantic meanings.
+After OCR, we used the quanteda package  [@Benoit2018] to remove punctuation, numbers, and stopwords, such as "the", "an", "a", due to their little or no semantic meanings.
 

### 2020-05-14 documents/paper.Rmd
work on method section

@@
 
-- something about basic steps
--- transform text into dfm, define dfm...
+- Basic steps for matrix construction
 
-- word embeddings (cite our readings)
--- a few sentences to introduce the reader to how WE have been used in other recent studies.
-this is a little bit background.
-'...recently used in many social science fields to...' tell the reader what was the finding of some key studies using WE that have inspired us.
+We transform text into a document-feature matrix (dfm), where rows represent PDFs and columns represent features, the terms extracted from PDFs, and values are frequencies of certain features for each PDFs.
+We selected the features that occur at least three times for later construction of a word co-occurrence matrix.
+In this article, we trained word embeddings with the GloVe model using th text2vec package [@Selivanov2016].
+The GloVe model calculates how often one word appear with the presence of another word to construct a co-occurrence matrix.
+This method involves setting an area defined by window in which we can detect the context terms with the target term in the middle position.
 
@@
 
--- then some more detail about the mechanics of the method, as we are using it. Say how what we did works in general.
+- temporal referencing plots (cite our readings)
+The word embeddings method is applied to explore temporal changes of vectors to present a historical trend.
+For example, @Garg2018 explore the historical trends in stereotypes and attitudes toward women and ethnic minorities in the 20th and 21st centuries in the US.
+They found both gender and ethnic occupation biases significantly correlated with the actual occupation participation rates.
+Those changes by negative words in the descriptions of genders and ethnic groups over time are associated with the women’s movement in the 1960s–1970s and Asian-American population growth in the 1960s and 1980s.
+Another study , @Kozlowski analyze the cultural dimensions of social class and how they evolve over the twentieth century in the US.
+They found the relationships between the cultural dimensions of class remain stable semantic structure over the century, but the individual words, such as employment, status, education, cultivation, on those dimensions shift over the decades.
+They also found affluence and status serve as cultural mediators between two clusters, education, cultivation, and morality on one side, and employment and ownership on the other side.
 
-In this article, we trained word embeddings with the GloVe model using th text2vec package [@Selivanov2016]. The GloVe model calculates ... creates the co-occurrence matrix.
+```{r temporal referencing plots}
+# Words per year
 
-- temporal referencing plots (cite our readings)
+library(tidyverse)
+
+# starting with https://quanteda.io/articles/quickstart.html
+library("quanteda")
+
+# get all txt files of abstracts
+abstract_txt_files <- 
+  fs::dir_ls(here::here("data/derived-data/"), 
+             recurse = TRUE, 
+             regexp = '\\.txt$')
+
+all_txts <- tolower(readtext::readtext(abstract_txt_files))
+names(all_txts) <- map_chr(names(all_txts), ~str_match_all(.x, "\\d{4}")[[1]][1,1])
+
+# word count checking
+all_txts_word_count_per_year <- 
+map_int(all_txts, 
+       stringi::stri_count_words) %>% 
+  stack %>% 
+  as_tibble() %>% 
+  rename(words = values,
+         year = ind) %>% 
+  mutate(year = as.numeric(as.character(year)))
+
+# count all words for each year
+all_txts_c <- corpus(all_txts)
+all_txts_c_summary <- 
+  summary(all_txts_c) %>%   
+  mutate(year = parse_number(Text))
+
+# make a dfm
+all_txts_c_dtm <- 
+  dfm(all_txts_c,
+      remove_punct = TRUE)
+
+# plot of all words in each year
+plot_all_words <- 
+  ggplot(data = all_txts_c_summary, 
+         aes(x = year , 
+             y = Tokens / 1000, 
+             group = 1)) +
+  geom_col() +
+  scale_x_continuous(labels = c(seq(1940, 2020, 2)), 
+                     breaks = seq(1940, 2020, 2),
+                     name = "Year") +
+  scale_y_continuous(name = "Total word count per year (x 1,000)",
+                     labels = scales::comma) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, 
+                                   vjust = 0.5))
+
+# Our key words over time
+
+keywords <- 
+  c("mechanism", 
+    "theory", 
+    "hypothesis", 
+    "model", 
+    "inference")
+
+dfm_keywords <- 
+  dfm_select(all_txts_c_dtm, 
+             pattern = keywords, 
+             selection = "keep")
+
+dfm_keywords_tbl <- 
+  convert(dfm_keywords, to = "data.frame") %>% 
+  pivot_longer(-document, 
+               names_to = "keyword",
+               values_to = "n") %>% 
+  mutate(year = parse_number(document))
+
+# compute proportion of all words per year
+dfm_keywords_tbl_prop <- 
+dfm_keywords_tbl %>% 
+  left_join(all_txts_c_summary) %>% 
+  mutate(prop = n / Tokens ) %>% 
+  group_by(keyword) %>% 
+  mutate(sum_the_word = sum(n)) %>% 
+  mutate(keyword_n = str_c(keyword, " (n = ", sum_the_word, ")"))
+
+# plot of keywords as a proportion of all words per year
+ggplot(data = dfm_keywords_tbl_prop, 
+       aes(x = year , 
+           y = prop)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(span = 0.475,
+              se = FALSE) +
+  facet_wrap( ~ keyword_n, 
+              ncol = 1,
+             scales = "free_y") +
+  scale_x_continuous(labels = c(seq(1960, 2020, 2)), 
+                     breaks = seq(1960, 2020, 2),
+                     name = "Year") +
+  scale_y_continuous(name = "Proportion of all words in all abstracts per year",
+                     labels = scales::comma) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, 
+                                   vjust = 0.5))
+
+```
 

### 2020-05-14 documents/paper.Rmd
editing text, polishing two phase word prop change plot

@@
 
-The study of disciplinary change traditionally relies on close reading of a large number of publications, manually sorting them into pre-defined categories and making connections between these groups and external influences [e.g.
+BM: The study of disciplinary change traditionally relies on close reading of a large number of publications, manually sorting them into pre-defined categories and making connections between these groups and external influences [e.g.
 @trigger1989history].
@@
 
-In this paper we present a dataset of all available abstracts of papers and posters presented at annual meetings of the Society of American Archaeology.
+BM: In this paper we present a dataset of all available abstracts of papers and posters presented at annual meetings of the Society of American Archaeology.
 We demonstrate the use of word embeddings, a commonly used computational tool in natural language processing (NLP) and machine learning, as a method to measure, quantify, and compare changes in the way archaeologists write cf.
@@
 
-Research into the history of archaeology is often organised into two major phases.
+BM: Research into the history of archaeology is often organised into two major phases.
 An early internalist phase, where the story of development of the field is told with limited reference to the influence of political, social, religious, and economic institutions.
@@
 
-Among historians of archaeology there are two major phases in  ‘critical history of archaeology’ in the 1980s had been characterized by the replacement of an
+BM: Among historians of archaeology there are two major phases in  ‘critical history of archaeology’ in the 1980s had been characterized by the replacement of an
 
@@
 - previous efforts at text analysis in archaeology (e.g. bibliometrics, citation networks)
+- new phase: we are now in the third phase  - computational text analysis
+- one promising new method is WE... successful applications of WE in other fields...
+
+- background to Word embeddings
+
+For example, @Garg2018 explored how stereotypes and attitudes toward women and ethnic minorities changed through 20th and 21st centuries in the US.
+<!-- how?-->  They found both gender and ethnic occupation biases significantly correlated with the actual occupation participation rates.
+Those changes by negative words in the descriptions of genders and ethnic groups over time are associated with the women’s movement in the 1960s–1970s and Asian-American population growth in the 1960s and 1980s.
+
+@Kozlowski analyzed the cultural dimensions of social class and how they evolved over the twentieth century in the US.
+<!-- how?--> They found the relationships between the cultural dimensions of class remain stable over the century, but the individual words, such as employment, status, education, cultivation, on those dimensions shift over the decades.
+<!-- for example ?--> They also found affluence and status serve as cultural mediators between two clusters <!-- of what?
+-->, education, cultivation, and morality on one side, and employment and ownership on the other side.
+<!-- I fear this is too abstract for our reader to grasp -->
 
 # Methods
-
-- source of the data
 
@@
 abstract_num <- sum(saa_data$number_of_abstracts, na.rm = TRUE)
-year_num <- length(which(!is.na(saa_data$number_of_abstracts)))
+year_num <- length(which(!is.na(saa_data$number_of_abstracts))) # 41 years vs 40
 ```
 
-The data for our study comes from the text of oral and poster presentation abstracts presented at the annual meetings of the Society of American Archaeology (SAA).
+LW: The data for our study comes from the text of abstracts of oral and poster presentations at the annual meetings of the Society of American Archaeology (SAA).
 The earliest meeting in our data is 1962, and we have some years missing, especially 1995-2003 where PDF files were not available to us.
@@
 
-The PDFs for the years 1962-1994 contain images of pages, and after 2005 the PDFs are searchable text.
+LW: The PDFs for the years 1962-1994 contain images of pages, and after 2005 the PDFs are searchable text.
 We converted the page images from PDF to PNG files using ImageMagick software, and then used optical character recognition to extract text from the PNG files with the Tesseract package [@Kay2007].
 We obtained an acceptable rate of 95-98% word-level OCR accuracy across the PDFs [@Holley2009].
-After OCR, we used the quanteda package  [@Benoit2018] to remove punctuation, numbers, and stopwords, such as "the", "an", "a", due to their little or no semantic meanings.
+After OCR, we removed punctuation, numbers, and stopwords, such as "the", "an", "a", due to their little or no semantic value for our research question, with the quanteda package [@Benoit2018].
 
-- Basic steps for matrix construction
+LW: We transformed the cleaned text into a document-feature matrix.
+This is a rectangular, tabular data structure where one row represents one PDF, and one column represents one word, also known as a feature.
+The cell values in this matrix are the frequencies of each feature for each PDF.
+We selected the features that occur at least three times across all PDFs for construction of a feature co-occurrence matrix.
+To create this matrix we calculated how often each feature appears with another.
+This method involves setting an area defined by window in which we can detect the context terms with the target term in the middle position.
+<!-- clarify and describe more fcm and window -->
 
-We transform text into a document-feature matrix (dfm), where rows represent PDFs and columns represent features, the terms extracted from PDFs, and values are frequencies of certain features for each PDFs.
-We selected the features that occur at least three times for later construction of a word co-occurrence matrix.
-In this article, we trained word embeddings with the GloVe model using th text2vec package [@Selivanov2016].
-The GloVe model calculates how often one word appear with the presence of another word to construct a co-occurrence matrix.
-This method involves setting an area defined by window in which we can detect the context terms with the target term in the middle position.
+We trained word embeddings with the GloVe model using the text2vec package [@Selivanov2016].
 
@@
 
-- temporal referencing plots (cite our readings)
-The word embeddings method is applied to explore temporal changes of vectors to present a historical trend.
-For example, @Garg2018 explore the historical trends in stereotypes and attitudes toward women and ethnic minorities in the 20th and 21st centuries in the US.
-They found both gender and ethnic occupation biases significantly correlated with the actual occupation participation rates.
-Those changes by negative words in the descriptions of genders and ethnic groups over time are associated with the women’s movement in the 1960s–1970s and Asian-American population growth in the 1960s and 1980s.
-Another study , @Kozlowski analyze the cultural dimensions of social class and how they evolve over the twentieth century in the US.
-They found the relationships between the cultural dimensions of class remain stable semantic structure over the century, but the individual words, such as employment, status, education, cultivation, on those dimensions shift over the decades.
-They also found affluence and status serve as cultural mediators between two clusters, education, cultivation, and morality on one side, and employment and ownership on the other side.
+LW: We investigated a historical trends in word meanings by visualizing the word embeddings generated by the GloVe method. ...
 

### 2020-05-14 code/0041-basic-word-stats-and-plots.R

@@ -9,6 +9,7 @@ dont_care <-
     'some', 'from', 'been', 'discuss', 'while',
     'over')
 
+# from tidytext
 df_tbl <- 
   convert(df, to = "data.frame") %>% 
   pivot_longer(-document, 
@@ -18,12 +19,19 @@ df_tbl <-
   filter(word %in% names(data_int_syllables)) %>% 
   filter(str_length(word) > 3) %>% 
   filter(!word %in% dont_care)
+
+df_tbl_yearly_totals <- 
+  df_tbl %>% 
+  group_by(year) %>% 
+  summarise(total_words = sum(n, na.rm = TRUE))
+  
   
 # compute proportion of all words per year
 df_tbl_prop <- 
   df_tbl %>% 
-  left_join(all_txts_c_summary) %>% 
-  mutate(prop = n / Tokens ) %>% 
+ # left_join(all_txts_c_summary) %>% 
+  left_join(df_tbl_yearly_totals) %>% 
+  mutate(prop = n / total_words ) %>% 
   filter(prop > 0) %>% 
   group_by(word) %>% 
   mutate(sum_the_word = sum(n)) %>% 
@@ -34,7 +42,9 @@ df_tbl_prop <-
 
 df_tbl_prop_two_groups <- 
   df_tbl_prop %>% 
-  mutate(year_group = ifelse(document < 2004, "early", "late"))
+  mutate(year_group = ifelse(document < 2004, 
+                             "early", 
+                             "late"))
 
 df_tbl_prop_two_groups_diff <- 
 df_tbl_prop_two_groups %>% 

### 2020-05-17 code/001-PDF-page-images-to-txt.R
add OCR for 1962 abstract

@@ -100,8 +100,51 @@ for(i in seq_len(length(abstracts_that_are_page_images_two_page_spread_paths))){
 }
 
 
+# Do OCR on PNG of single pages
 
+abstracts_that_are_page_images_single <- 
+  abstracts_tally %>% 
+  filter(single_or_double_page_spread == "single")
 
+# OCR abstracts that are images of single pages
+  
+# select PDF document to work on 
+year_single <- abstracts_that_are_page_images_single$year
+page_start_single <- abstracts_that_are_page_images_single$abstract_start_page
+page_end_single <- abstracts_that_are_page_images_single$abstract_end_page
+  
+# create directory to hold results
+wd <- getwd()
+dir_create(paste0(here::here("data/derived-data/abstracts-ocr/"), year_single, "/pngs"))
+setwd(paste0(here::here("data/derived-data/abstracts-ocr/"), year_single, "/pngs"))
+  
+# convert PDF to PNG
+pdf_convert_files_single <- 
+  pdf_file_names %>% 
+  str_subset(as.character(year_single)) %>% 
+  pdftools::pdf_convert(pages = page_start_single:page_end_single, 
+                          dpi = 600,
+  )
+  
+setwd(wd)
+
+# Do OCR on PNG of single pages
+
+pngs_single_path <- paste0("data/derived-data/abstracts-ocr/", year_single, "/pngs")
+png_single_files <- list.files(pngs_single_path, full.names = TRUE)
+
+text <- vector("list", length = length(png_single_files))
+for(i in seq_len(length(png_single_files))){
+  text [[i]] <- tesseract::ocr(png_single_files[i]) 
+}
+
+text_c <- paste0(text, collapse = ", ") # combine into a single element
+dir_create(paste0("data/derived-data/abstracts-ocr/", year_single, "/txt_from_split_pages"))
+write_lines(text_c, 
+            paste0("data/derived-data/abstracts-ocr/", year_single, 
+                   "/txt_from_split_pages/abstract-text-", 
+                   year_single, 
+                   ".txt"))
 
 
 # convert PDF to text

### 2020-05-17 documents/paper.Rmd
add in-line code for word counts

@@
 saa_data <- readxl::read_excel(here::here("data", "raw-data", "saa-abstracts-tally.xlsx"))
+
+# basic stats
+library(tidyverse)
 abstract_num <- sum(saa_data$number_of_abstracts, na.rm = TRUE)
-year_num <- length(which(!is.na(saa_data$number_of_abstracts))) # 41 years vs 40
+year_num <- length(which(!is.na(saa_data$number_of_abstracts))) # 41 years
+
+# get all txt files of abstracts
+abstract_txt_files <- 
+  fs::dir_ls(here::here("data/derived-data/"), 
+             recurse = TRUE, 
+             regexp = '\\.txt$')
+all_txts <- tolower(readtext::readtext(abstract_txt_files))
+
+all_word_counts <- sum(stringi::stri_count_words(all_txts))
 ```
@@
 The earliest meeting in our data is 1962, and we have some years missing, especially 1995-2003 where PDF files were not available to us.
-The total number of abstracts is `r prettyNum(abstract_num, big.mark = ",")` (xxx words) from `r year_num` annual meetings.
+The total number of abstracts is `r prettyNum(abstract_num, big.mark = ",")` (`r prettyNum(all_word_counts, big.mark = ",")` words) from `r year_num` annual meetings.
 We obtained PDF files of annual programs from https://www.saa.org/annual-meeting/annual-meeting-archives/program-archives.

### 2020-05-19 documents/paper.Rmd
remove code chunk

@@
 
-```{r temporal referencing plots}
-# Words per year
-
-library(tidyverse)
-
-# starting with https://quanteda.io/articles/quickstart.html
-library("quanteda")
-
-# get all txt files of abstracts
-abstract_txt_files <- 
-  fs::dir_ls(here::here("data/derived-data/"), 
-             recurse = TRUE, 
-             regexp = '\\.txt$')
-
-all_txts <- tolower(readtext::readtext(abstract_txt_files))
-names(all_txts) <- map_chr(names(all_txts), ~str_match_all(.x, "\\d{4}")[[1]][1,1])
-
-# word count checking
-all_txts_word_count_per_year <- 
-map_int(all_txts, 
-       stringi::stri_count_words) %>% 
-  stack %>% 
-  as_tibble() %>% 
-  rename(words = values,
-         year = ind) %>% 
-  mutate(year = as.numeric(as.character(year)))
-
-# count all words for each year
-all_txts_c <- corpus(all_txts)
-all_txts_c_summary <- 
-  summary(all_txts_c) %>%   
-  mutate(year = parse_number(Text))
-
-# make a dfm
-all_txts_c_dtm <- 
-  dfm(all_txts_c,
-      remove_punct = TRUE)
-
-# plot of all words in each year
-plot_all_words <- 
-  ggplot(data = all_txts_c_summary, 
-         aes(x = year , 
-             y = Tokens / 1000, 
-             group = 1)) +
-  geom_col() +
-  scale_x_continuous(labels = c(seq(1940, 2020, 2)), 
-                     breaks = seq(1940, 2020, 2),
-                     name = "Year") +
-  scale_y_continuous(name = "Total word count per year (x 1,000)",
-                     labels = scales::comma) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, 
-                                   vjust = 0.5))
-
-# Our key words over time
-
-keywords <- 
-  c("mechanism", 
-    "theory", 
-    "hypothesis", 
-    "model", 
-    "inference")
-
-dfm_keywords <- 
-  dfm_select(all_txts_c_dtm, 
-             pattern = keywords, 
-             selection = "keep")
-
-dfm_keywords_tbl <- 
-  convert(dfm_keywords, to = "data.frame") %>% 
-  pivot_longer(-document, 
-               names_to = "keyword",
-               values_to = "n") %>% 
-  mutate(year = parse_number(document))
-
-# compute proportion of all words per year
-dfm_keywords_tbl_prop <- 
-dfm_keywords_tbl %>% 
-  left_join(all_txts_c_summary) %>% 
-  mutate(prop = n / Tokens ) %>% 
-  group_by(keyword) %>% 
-  mutate(sum_the_word = sum(n)) %>% 
-  mutate(keyword_n = str_c(keyword, " (n = ", sum_the_word, ")"))
-
-# plot of keywords as a proportion of all words per year
-ggplot(data = dfm_keywords_tbl_prop, 
-       aes(x = year , 
-           y = prop)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(span = 0.475,
-              se = FALSE) +
-  facet_wrap( ~ keyword_n, 
-              ncol = 1,
-             scales = "free_y") +
-  scale_x_continuous(labels = c(seq(1960, 2020, 2)), 
-                     breaks = seq(1960, 2020, 2),
-                     name = "Year") +
-  scale_y_continuous(name = "Proportion of all words in all abstracts per year",
-                     labels = scales::comma) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, 
-                                   vjust = 0.5))
-
-```
-
 # Results

### 2020-05-21 documents/paper.Rmd
code tidy in Rmd

@@
 saa_data <- readxl::read_excel(here::here("data", "raw-data", "saa-abstracts-tally.xlsx"))
-
-# basic stats
-library(tidyverse)
 abstract_num <- sum(saa_data$number_of_abstracts, na.rm = TRUE)
-year_num <- length(which(!is.na(saa_data$number_of_abstracts))) # 41 years
-
-# get all txt files of abstracts
-abstract_txt_files <- 
-  fs::dir_ls(here::here("data/derived-data/"), 
-             recurse = TRUE, 
-             regexp = '\\.txt$')
-all_txts <- tolower(readtext::readtext(abstract_txt_files))
-
-all_word_counts <- sum(stringi::stri_count_words(all_txts))
+year_num <- length(which(!is.na(saa_data$number_of_abstracts))) # 41 years vs 40
 ```
@@
 The earliest meeting in our data is 1962, and we have some years missing, especially 1995-2003 where PDF files were not available to us.
-The total number of abstracts is `r prettyNum(abstract_num, big.mark = ",")` (`r prettyNum(all_word_counts, big.mark = ",")` words) from `r year_num` annual meetings.
+The total number of abstracts is `r prettyNum(abstract_num, big.mark = ",")` (xxx words) from `r year_num` annual meetings.
 We obtained PDF files of annual programs from https://www.saa.org/annual-meeting/annual-meeting-archives/program-archives.
@@
 
+```{r temporal-referencing-plots, cache.extra=tools::md5sum(here::here("006-testing-word-2-vec-temporal-referencing-by-period.R"))}
+source(here::here("code/006-testing-word-2-vec-temporal-referencing-by-period.R"))
+```
+
 # Results

### 2020-05-21 code/006-testing-word-2-vec-temporal-referencing-by-period.R

@@ -24,9 +24,12 @@ all_txts_c_dtm <-
 # investigate the target feature a bit
 
 # we can change these
-related_words <- c("mechanism", "mechanisms")
+related_words <- c("theory", "theories", "theoretical")
 map(related_words, ~colSums(all_txts_c_dtm[, .x ]))
 #  so use the most common one
+
+# mechanisms
+# theory
 
 
 # https://www.aclweb.org/anthology/P19-1044.pdf is our inspiration 
@@ -93,9 +96,9 @@ all_txts_c_fcm <- fcm(all_txts_c_dtm_toks2,
 
 library("text2vec")
 
-glove <- GlobalVectors$new(rank = 50, x_max = 10)
+glove <- GlobalVectors$new(rank = 50, x_max = 20)
 all_txts_main <- glove$fit_transform(all_txts_c_fcm, 
-                               n_iter = 10,
+                               n_iter = 20,
                                convergence_tol = -1, 
                                n_threads = 8)
 
@@ -135,7 +138,7 @@ map(time_specific_token,
   names() %>% 
   unique()
 
-# keep only the tokens found in an English dictionary
+# keep only the tokens found in an English syllables dictionary
 myTokens <- 
   featnames(dfm(tokens_select(tokens(similar_words), 
                               names(data_int_syllables))))
@@ -172,7 +175,7 @@ tsne_plot <-
             aes(x = V1, 
                 y = V2, 
                 label = word), 
-            size = 3) +
+            size = 2) +
   geom_segment(data = tsne_plot_target_features, 
                aes(x = V1, 
                    y = V2,
@@ -185,7 +188,7 @@ tsne_plot <-
             aes(x = V1, 
                 y = V2,
                 label = word),
-            size = 3,
+            size = 2,
             colour = "red",
             bg.color = "white", 
             bg.r = 0.15 ) +
@@ -199,3 +202,5 @@ tsne_plot +
                  " year groupings ")) +
   labs(caption  = "Data and code online at https://github.com/benmarwick/saa-meeting-abstracts")
 
+ggsave(here::here(str_glue("figures/tr-plot-target-word-is-{target_feature}-{year_interval}-by-year-intervals.png")))
+  

### 2020-05-21 documents/paper.Rmd
update background and method section

--- 2020-05-19 documents/paper.Rmd
@@
 
-For example, @Garg2018 explored how stereotypes and attitudes toward women and ethnic minorities changed through 20th and 21st centuries in the US.
-<!-- how?-->  They found both gender and ethnic occupation biases significantly correlated with the actual occupation participation rates.
-Those changes by negative words in the descriptions of genders and ethnic groups over time are associated with the women’s movement in the 1960s–1970s and Asian-American population growth in the 1960s and 1980s.
+LW: Word embeddings are an effective way to analyze the temporal changes of any target word, which is represented by a group of words sharing close semantic relations with it.
+Word embeddings needs to be trained on a corpus, a collection of texts, for later calculation of the relations between words according to their distance in texts.
+This method allows us to explore the changing meaning of words of interest accompanied with their social or cultural background to understand the historical trend.
+For example, @Garg2018 explored changes in gender stereotypes and attitudes toward ethnic minorities in the 20th and 21st centuries in the US.
+They created word lists to represent gender and ethnicity as target words, and neutral words, such as adjectives and jobs, serving as indicators for stereotypes and attitudes.
+With word embeddings, they compared the association between each group and neutral words, where the strength of association is calculated by computing average embedding distance.
+They found there is a job bias for both gender and ethnic minority, such as and this bias is significantly correlated with the actual participation rates.
+In addition, there is bias towards  decreases of bias reflected by adjectives for ethnic minority and gender are associated with the women’s movement in the 1960s–1970s and Asian-American population growth in the 1960s and 1980s.
 
-@Kozlowski analyzed the cultural dimensions of social class and how they evolved over the twentieth century in the US.
+Another example shows how we can use it to explore changes in the definition of social class.
+@Kozlowski analyzed the cultural dimensions of social class over the 20th century in the US.
 <!-- how?--> They found the relationships between the cultural dimensions of class remain stable over the century, but the individual words, such as employment, status, education, cultivation, on those dimensions shift over the decades.
@@
 To create this matrix we calculated how often each feature appears with another.
-This method involves setting an area defined by window in which we can detect the context terms with the target term in the middle position.
-<!-- clarify and describe more fcm and window -->
+The co-occurrences is measured within a defined range, called a window, with the target feature in the middle position and context features next to it.
+Here we set the range to five that means five words before and after the target feature as the framework for calculation of co-occurrence counts.
 
-We trained word embeddings with the GloVe model using the text2vec package [@Selivanov2016].
+LW: We trained the word embeddings on the co-occurrence matrix by GloVe model using the text2vec package [@Selivanov2016].
+GloVe (Global Vectors) is a learning model trained on global information, in our case is the co-occurrence matrix from abtract corpus, to capture word vectors, a group of words, as a representation of a target word.
+The underlying principle is that GloVe algorithm converts co-occurrence counts to probabilities ratios for words of interest.
 
 <!-- There are two major word-level pretrained word embeddings, Word2Vec and Glove according to different methods used for training [@Rong2014; @Pennington2014]. -->
-
-LW: We investigated a historical trends in word meanings by visualizing the word embeddings generated by the GloVe method. ...
 

### 2020-05-21 documents/paper.Rmd
udpate background and add para about GloVe

@@
 In addition, there is bias towards  decreases of bias reflected by adjectives for ethnic minority and gender are associated with the women’s movement in the 1960s–1970s and Asian-American population growth in the 1960s and 1980s.
+This study demonstrates a successful application of word embeddings to detect social bias and correlate the findings with historical events.
 
-Another example shows how we can use it to explore changes in the definition of social class.
-@Kozlowski analyzed the cultural dimensions of social class over the 20th century in the US.
-<!-- how?--> They found the relationships between the cultural dimensions of class remain stable over the century, but the individual words, such as employment, status, education, cultivation, on those dimensions shift over the decades.
-<!-- for example ?--> They also found affluence and status serve as cultural mediators between two clusters <!-- of what?
--->, education, cultivation, and morality on one side, and employment and ownership on the other side.
-<!-- I fear this is too abstract for our reader to grasp -->
+LW: Another example shows how word embeddings is effective by exploring the historical view on shared understandings of social classes in the 20th century in the US.
+@Kozlowski first compared the cultural dimensions created by word embeddings with a defined word lists showing cultural differences, such as rich-poor.
+The high association suggests that word embedding is a useful way to investigate people's views about social classes.
+They explored the relationship between dimensions of class identified in sociological theory, including morality, status, education, cultivation, gender, and employment.
+They found there is a high correlation between dimensions of education, cultivation, and morality, which corresponds to the owner/worker relation defined by classic Marxian.
+In addition, they found a stable relationships between different cultural dimensions of class over the century even there was a major shifts in the condition of economy, industry, and employment in the present day.
+This case study indicates the robustness of word embeddings as a method to examine the historical trend on social classes.
 
@@
 
-LW: We trained the word embeddings on the co-occurrence matrix by GloVe model using the text2vec package [@Selivanov2016].
-GloVe (Global Vectors) is a learning model trained on global information, in our case is the co-occurrence matrix from abtract corpus, to capture word vectors, a group of words, as a representation of a target word.
-The underlying principle is that GloVe algorithm converts co-occurrence counts to probabilities ratios for words of interest.
+LW: We trained the word embeddings on the co-occurrence matrix by GloVe model using the text2vec package [@Pennington2014; @Selivanov2016].
+GloVe (Global Vectors) is a learning model that creates word vectors to represent semantic regularity and capture relationships between words [@Pennington2014; @Garg2018].
+A word vector can be thought as a point with direction in a space, and it has multiple dimension associated with other word vectors.
+The distance between word vectors is measured by their various dimensions of difference, representing their semantic meanings.
+The difference is computed based on their co-occurrence probabilities converted from  co-occurrence matrix.
+This idea is similar to the concept of Principal Component Analysis (PCA).
+A dimension consists of words with its scale and direction can be analogous to a dimension generated by Principal Component Analysis.
+The words located on the two ends present the largest semantic distance represented by values.
+The more dimensions, the more accurate semantic meanings can be captured.
+In our case, we use 50 dimensions.
+
+LW: We then visualize the values across different dimensions to explore the historical trend of word meanings related to disciplinary change in archaeology.
+We created a word list, including theory, inference, mechanism, explanation, hypothesis, to explore how they changes in their meaning or use since 1962.
+We compare the word embeddings at a interval of ten years.
 

### 2020-05-21 documents/paper.Rmd
simplify TR code into custom functions, run on a bunch of key words and make plots

--- 2020-05-21 documents/paper.Rmd
@@
 
-# Results
-
 - basic details (e.g. word and abstract counts)

### 2020-05-21 code/006-temp-ref-functions.R

@@ -1,0 +1,237 @@
+
+#-----------------------------------------------------------
+read_in_the_abstracts_fn <- function(){
+
+library(tidyverse)
+
+# starting with https://quanteda.io/articles/quickstart.html
+library("quanteda")
+
+# get all txt files of abstracts
+abstract_txt_files <- 
+fs::dir_ls(here::here("data/derived-data/"), 
+           recurse = TRUE, 
+           regexp = '\\.txt$')
+
+all_txts <- tolower(readtext::readtext(abstract_txt_files))
+names(all_txts) <- map_chr(names(all_txts), ~str_match_all(.x, "\\d{4}")[[1]][1,1])
+
+all_txts_c <- corpus(all_txts)
+
+# make a dfm, removing stopwords and maybe applying stemming
+all_txts_c_dtm <- 
+  dfm(all_txts_c,
+      remove = stopwords("english"),
+      #stem = TRUE, 
+      remove_punct = TRUE)
+
+return(list(all_txts = all_txts,
+            all_txts_c_dtm = all_txts_c_dtm))
+
+}
+
+
+#-----------------------------------------------------------
+
+insert_time_specific_token_fn <- function(all_txts = all_txts,
+                                          target_feature = "mechanisms",
+                                          year_interval = 5){
+
+# https://www.aclweb.org/anthology/P19-1044.pdf is our inspiration 
+# in each text we want to replace the target feature (word) with a tagged feature,
+# a time-specific token target word w ∈ Ct with a time-specific token wt
+# E.g., in the corpus for 1920 we replace each occurrence of computer with the string computer1920
+
+year <- names(all_txts)
+
+floor_to_interval <-  function(value, interval){ return(value - value %% interval) }
+
+# year_interval <- 5 # we can change and explore
+interval <- floor_to_interval(as.numeric(year), year_interval)
+# target_feature <- "mechanisms" # we can change this
+
+time_specific_token <- 
+  paste0(target_feature, "_", interval, "_", interval + year_interval, "")
+
+all_txts_updated <- vector("character", length = length(all_txts))
+for(i in seq_len(length(all_txts))){
+  
+  tmp1 <- all_txts[i]
+  tmp2 <- str_replace_all(tmp1, "-", "") 
+  tmp2 <- str_replace_all(tmp2, "\n", " ")
+  tmp3 <- time_specific_token[i]
[…]

### 2020-05-21 code/006-temp-ref-scripts.R

@@ -1,0 +1,105 @@
+
+
+#-----------------------------------------------------------
+# read in the abstracts text files
+# takes a long time!
+read_in_the_abstracts_data <- 
+  read_in_the_abstracts_fn()
+
+# save it so we don't have to repeat this each time
+saveRDS(read_in_the_abstracts_data, 
+        here::here("data/derived-data/read_in_the_abstracts_data.rds"))
+#-----------------------------------------------------------
+
+# investigate the target feature a bit
+
+# we can explore these
+related_words <- c("theory", "theories", "theoretical")
+map(related_words, ~colSums(read_in_the_abstracts_data$all_txts_c_dtm[, .x ]))
+#  so use the most common one, eg.:
+
+# model
+# theory
+# mechanisms
+# hypothesis
+# explanations
+# inferences
+
+# narrative'
+# interpretation
+# understanding
+# agency
+# practice
+
+# set the word here that we use in the later functions:
+target_feature <-  "theory"
+year_interval <-  5
+
+insert_generate_plot_fn <- 
+  function(target_feature, 
+           year_interval, 
+           read_in_the_abstracts_data){
+
+# Now we insert the time-specific token, like "theory_1975_1980"
+
+#-----------------------------------------------------------
+# insert time specific token in the dfm
+# takes a few moments 
+insert_time_specific_token_data <- 
+  insert_time_specific_token_fn(read_in_the_abstracts_data$all_txts,
+                                target_feature,
+                                year_interval)
+#-----------------------------------------------------------
+
+# Now we can generate the word embeddings for our target feature:
+
+#-----------------------------------------------------------
+# generate word embeddings
+# takes a long time!
+generate_embedding_matrix_data <- 
+  generate_embedding_matrix_fn(insert_time_specific_token_data$all_txts_c_fcm,
+                               insert_time_specific_token_data$time_specific_token,
+                               n_similar_words = 50)
+#-----------------------------------------------------------
+
+# Now we can draw the plot:
+
+#-----------------------------------------------------------
+# draw a plot
+# pretty fast
+temporal_referencing_plot_output <- 
+  temporal_referencing_plot_fn(generate_embedding_matrix_data$embedding_matrix,
+                               generate_embedding_matrix_data$myTokens,
+                               time_specific_token = insert_time_specific_token_data$time_specific_token,
+                               target_feature,
[…]

### 2020-05-21 code/006-testing-word-2-vec-temporal-referencing-by-period.R

@@ -1,206 +1,0 @@
-library(tidyverse)
-
-# starting with https://quanteda.io/articles/quickstart.html
-library("quanteda")
-
-# get all txt files of abstracts
-abstract_txt_files <- 
-fs::dir_ls(here::here("data/derived-data/"), 
-           recurse = TRUE, 
-           regexp = '\\.txt$')
-
-all_txts <- tolower(readtext::readtext(abstract_txt_files))
-names(all_txts) <- map_chr(names(all_txts), ~str_match_all(.x, "\\d{4}")[[1]][1,1])
-
-all_txts_c <- corpus(all_txts)
-
-# make a dfm, removing stopwords and maybe applying stemming
-all_txts_c_dtm <- 
-  dfm(all_txts_c,
-      remove = stopwords("english"),
-      #stem = TRUE, 
-      remove_punct = TRUE)
-
-# investigate the target feature a bit
-
-# we can change these
-related_words <- c("theory", "theories", "theoretical")
-map(related_words, ~colSums(all_txts_c_dtm[, .x ]))
-#  so use the most common one
-
-# mechanisms
-# theory
-
-
-# https://www.aclweb.org/anthology/P19-1044.pdf is our inspiration 
-# in each text we want to replace the target feature (word) with a tagged feature,
-# a time-specific token target word w ∈ Ct with a time-specific token wt
-# E.g., in the corpus for 1920 we replace each occurrence of computer with the string computer1920
-
-year <- names(all_txts)
-
-floor_to_interval <-  function(value, interval){ return(value - value %% interval) }
-
-year_interval <- 5 # we can change and explore
-interval <- floor_to_interval(as.numeric(year), year_interval)
-target_feature <- "mechanisms" # we can change this
-
-time_specific_token <- 
-  paste0(target_feature, "_", interval, "_", interval + year_interval, "")
-
-all_txts_updated <- vector("character", length = length(all_txts))
-for(i in seq_len(length(all_txts))){
-  
-  tmp1 <- all_txts[i]
-  tmp2 <- str_replace_all(tmp1, "-", "") 
-  tmp2 <- str_replace_all(tmp2, "\n", " ")
-  tmp3 <- time_specific_token[i]
-  tmp4 <- str_c("\\b", target_feature, "\\b", collapse="|")
-  tmp5 <- str_replace_all(tmp2, tmp4, tmp3)
-  all_txts_updated[[i]] <- tmp5
-
-}
-
-all_txts_updated_c <- corpus(all_txts_updated)
-
[…]

### 2020-05-21 documents/paper.Rmd
edits to text

@@
 
-For example, @Garg2018 explored how stereotypes and attitudes toward women and ethnic minorities changed through 20th and 21st centuries in the US.
-<!-- how?-->  They found both gender and ethnic occupation biases significantly correlated with the actual occupation participation rates.
-Those changes by negative words in the descriptions of genders and ethnic groups over time are associated with the women’s movement in the 1960s–1970s and Asian-American population growth in the 1960s and 1980s.
+LW: Word embeddings are an effective way to analyze the temporal changes of any target word.
+A word embedding is a set of numbers that characterizes the meaning a word.
+Word embeddings are generated on a collection of texts, and can be used for calculation of the semantic relationships between words according to their distance in the texts.
+We can see how the meaning of a word shifts over time by visualizing the word embeddings and observing how the word moves through different spaces occupied by other words over time.
+When a word is in close proximity to other words, they have a high degree of semantic relatedness.
+This method allows for the exploration of the changing meaning of words in different contexts.
+For example, @Garg2018 explored changes in gender stereotypes and attitudes toward ethnic minorities in the 20th and 21st centuries in the US.
+They created custom lists of target words representing gender and ethnicity, and neutral words, such as adjectives and jobs, serving as indicators for stereotypes and attitudes.
+<!--where did their word embeddings come from, and what algorithm did they use?
+--> With word embeddings, they compared the association between each group <!--what is the group?
+we need to mention the groups before this sentence --> and neutral words, where the strength of association is calculated by computing average embedding distance.
+They found there is a job bias for both gender and ethnic minority, <!-- we really need a concrete example here --> this bias is significantly correlated with the actual job participation rates.
+In addition, there is a decreases in bias reflected by adjectives for ethnic minority and gender are associated with the women’s movement in the 1960s–1970s and Asian-American population growth in the 1960s and 1980s <!--"Another bias they identified was ...
+give the reader a concrete result from the paper  -->.
+This study demonstrates a successful application of word embeddings to detect social bias <!-- can we re-prhase to be about changing meanings of words?
+--> and correlate the findings with historical events.
 
-@Kozlowski analyzed the cultural dimensions of social class and how they evolved over the twentieth century in the US.
-<!-- how?--> They found the relationships between the cultural dimensions of class remain stable over the century, but the individual words, such as employment, status, education, cultivation, on those dimensions shift over the decades.
-<!-- for example ?--> They also found affluence and status serve as cultural mediators between two clusters <!-- of what?
--->, education, cultivation, and morality on one side, and employment and ownership on the other side.
-<!-- I fear this is too abstract for our reader to grasp -->
+LW: Another example shows how word embeddings are effective <!-- Q: for what?
+A: for understanding semantic change over time --> by exploring the historical view on shared understandings of social classes in the 20th century in the US.
+@Kozlowski first compared the cultural dimensions created by word embeddings <!-- generated word embeddings from what data?
+and what algorithm did they use?
+--> with a defined word list reflecting socio-economic differences, such as 'rich', 'poor' and so on.
+The high association <!-- of what?
+between what?
+--> suggests that word embedding is a useful way to investigate people's views about social classes.
+They explored the relationship between dimensions of class identified in sociological theory, including morality, status, education, cultivation, gender, and employment.
+They found there is a high correlation <!-- of what?
+spell it out if this is about the word vectors, say --> between dimensions of education, cultivation, and morality, which corresponds to the owner/worker relation <!-- what relation?
+spell out more --> defined by classic Marxism.
+In addition, they found a stable relationship between different cultural dimensions of class over the century even though there was a major shift in the condition of economy, industry, and employment in the present day.
+This case study indicates the robustness of word embeddings as a method to examine historical trends of social classes.
 
@@
 To create this matrix we calculated how often each feature appears with another.
-This method involves setting an area defined by window in which we can detect the context terms with the target term in the middle position.
-<!-- clarify and describe more fcm and window -->
+The co-occurrences are measured within a defined range of surrounding words, known as a window, with the target feature in the middle position and context features before and after it.
+Here we set the range to five, meaning five words before and five words after the target feature, as our framework for calculating the co-occurrence counts.
 
-We trained word embeddings with the GloVe model using the text2vec package [@Selivanov2016].
+LW: We generated the word embeddings on the co-occurrence matrix using the GloVe algorithm from the text2vec package for R [@Pennington2014; @Selivanov2016].
+GloVe (Global Vectors) is a unsupervised learning model that creates word vectors to represent semantic regularity and capture relationships between words [@Pennington2014; @Garg2018].
+This idea is similar to the concept of Principal Component Analysis (PCA) that reduces a multi-dimensional dataset into a small number of new dimensions that usually capture most of the variation in the data.
+In this case, a word vector is similar to a principal component because it expresses the meaning of the word into a small set of numeric values.
+The more values in this set, or the more dimensions, the more accurately semantic meanings of a word can be captured.
+In our case, we use 50 dimensions.
+One of the most interesting qualities of word vectors is that simple arithmetic operations can reveal semantic relationships between words [@mikolov2013linguistic].
+For example if we subtract the word vector values for 'man' from the word vector values for 'king', and add the word vector values for 'woman', the resulting vector will be very close to that of 'queen', such that we can literally compute queen = king - man + woman.
+
+<!-- one paragraph of plain and simple description of how Glove computes word vectors from co-oc matrix -->
+
+LW: We then visualize the values across different dimensions to explore the historical trend of word meanings related to disciplinary change in archaeology.
+
+<!-- one paragraph of explaining how did we choose these words -->
+
+We created a word list, including theory, inference, mechanism, explanation, hypothesis, to explore how they changes in their meaning or use since 1962.
+We compare the word embeddings at a interval of ten years.
 
 <!-- There are two major word-level pretrained word embeddings, Word2Vec and Glove according to different methods used for training [@Rong2014; @Pennington2014]. -->
-
-LW: We investigated a historical trends in word meanings by visualizing the word embeddings generated by the GloVe method. ...
 

### 2020-05-22 documents/paper.Rmd
editing background & methods

@@
 
-LW: Word embeddings are an effective way to analyze the temporal changes of any target word.
-A word embedding is a set of numbers that characterizes the meaning a word.
-Word embeddings are generated on a collection of texts, and can be used for calculation of the semantic relationships between words according to their distance in the texts.
-We can see how the meaning of a word shifts over time by visualizing the word embeddings and observing how the word moves through different spaces occupied by other words over time.
-When a word is in close proximity to other words, they have a high degree of semantic relatedness.
-This method allows for the exploration of the changing meaning of words in different contexts.
+LW & BM: Word embeddings are increasingly being used by researchers in novel ways and have become popular in fields such as digital humanities and computational social science.
+A word embedding is a set of numbers representing a word that generated from a collection of texts.
+A word embedding characterizes the meaning a word and can be used to reveal latent semantic relationships between pairs of words, often measured by computing the cosine similarity of two embeddings [@antoniak2018evaluating].
+The general idea is similar to the concept of Principal Component Analysis (PCA) that reduces a multi-dimensional dataset into a small number of new dimensions that typically capture most of the variation in the data.
+In this case, a word vector is similar to a principal component because it expresses the meaning of the word into a small set of numeric values (e.g.
+typically 50-250 values).
+This set is known as the word embedding or word vector, and we visualize vectors on a plot, the distance between two words estimates the similarity between the two words.
+One of the most interesting and useful qualities of word vectors is that simple arithmetic operations can reveal semantic relationships between words [@mikolov2013linguistic].
+For example if we subtract the word vector values for 'man' from the word vector values for 'king', and add the word vector values for 'woman', the resulting vector will be very close to that of 'queen', such that we can literally compute 'king - man + woman' to get a vector that is approximately the same as 'queen'.
+
+BM: Word embeddings have been used as evidence in studies of historical change in language and culture [@phillips2017intrinsic; @heuser2017word; @kulkarni2015statistically].
+Semantic relations between words in a corpus are captured by the geometry of the vectors, so vectors can be compared with each other to find a word’s nearest neighbors [@giorcelli2019does].
+Vectors can be compared across periods to examine the degree of change a word underwent from one-time interval to the next.
+For example, @hamilton2016diachronic use word embeddings to study semantic change in English over the last 200 years.
+They visualise how the word 'gay' shifted from meaning 'cheerful' or 'frolicsome' to referring to homosexuality, and the change in 'broadcast' from referring to 'casting out seeds' to 'transmitting signals', with the rise of television and radio in the early 20th century.
+@yao2018dynamic generate word embeddings on 99,872 articles from the New York Times, published between January 1990 and July 2016 and their results show major semantic shifts in their words of interest.
+For example, the word embeddings for the US president names, Obama and Trump, shift from their pre-presidential lives (Obama as a civil rights attorney and university professor; Trump as a real estate developer and TV personality) to the political sphere.
+
+LW:  Word embeddings  have also been used to track demographic and occupational social shifts and gender stereotypes.
 For example, @Garg2018 explored changes in gender stereotypes and attitudes toward ethnic minorities in the 20th and 21st centuries in the US.
@@
 GloVe (Global Vectors) is a unsupervised learning model that creates word vectors to represent semantic regularity and capture relationships between words [@Pennington2014; @Garg2018].
-This idea is similar to the concept of Principal Component Analysis (PCA) that reduces a multi-dimensional dataset into a small number of new dimensions that usually capture most of the variation in the data.
-In this case, a word vector is similar to a principal component because it expresses the meaning of the word into a small set of numeric values.
-The more values in this set, or the more dimensions, the more accurately semantic meanings of a word can be captured.
-In our case, we use 50 dimensions.
-One of the most interesting qualities of word vectors is that simple arithmetic operations can reveal semantic relationships between words [@mikolov2013linguistic].
-For example if we subtract the word vector values for 'man' from the word vector values for 'king', and add the word vector values for 'woman', the resulting vector will be very close to that of 'queen', such that we can literally compute queen = king - man + woman.
 
@@
 
-LW: We then visualize the values across different dimensions to explore the historical trend of word meanings related to disciplinary change in archaeology.
+BM: One of the key practical challenges of studying temporal change in word embeddings is aligning the vectors representing the same words across multiple time periods.
+This is because the stochastic nature of word embeddings algorithms makes them generate representations that always have different coordinate systems and thus one set of embeddings are are not directly comparable to another set generated on a different corpus (i.e.
+from a different time period).
+Previous work generally follows a two-step pattern: first compute word embeddings in each time slice separately, then find a way to align the word embeddings across time slices.
+@kim2014temporal generated word embeddings for each time period by initializing them with the embeddings from the previous period.
+@kulkarni2015statistically compute a linear transformation of words between any two time slices by solving a n-dimensional least squares problem of k nearest neighbor words (where n is the embedding dimension).
+@zhang2016past similarly use a linear transformation between a base and target time slices, and compute the
+linear transformation using anchor words that does not change meaning between the two time slices (requiringprior knowledge of words that are static).
+@hamilton2016diachronic solves a n-dimensional Procrustes problem between every two adjacent time slices to perform the alignment.
+
+We then visualize the values across different dimensions to explore the historical trend of word meanings related to disciplinary change in archaeology.
 

### 2020-05-22 code/006-temp-ref-functions.R

@@ -108,11 +108,12 @@ return(list(all_txts_c_fcm = all_txts_c_fcm,
  
 generate_embedding_matrix_fn <- function(all_txts_c_fcm = all_txts_c_fcm,
                                          time_specific_token = time_specific_token,
-                                         n_similar_words = 50){
+                                         n_similar_words = 50,
+                                         rank = 100){
 
 library("text2vec")
 
-glove <- GlobalVectors$new(rank = 50, x_max = 20)
+glove <- GlobalVectors$new(rank = rank, x_max = 20)
 all_txts_main <- glove$fit_transform(all_txts_c_fcm, 
                                n_iter = 20,
                                convergence_tol = -1, 

### 2020-05-22 code/006-temp-ref-scripts.R

@@ -9,12 +9,18 @@ read_in_the_abstracts_data <-
 # save it so we don't have to repeat this each time
 saveRDS(read_in_the_abstracts_data, 
         here::here("data/derived-data/read_in_the_abstracts_data.rds"))
+
+# read it in to save time
+read_in_the_abstracts_data <- 
+  readRDS(here::here("data/derived-data/read_in_the_abstracts_data.rds"))
 #-----------------------------------------------------------
 
 # investigate the target feature a bit
 
 # we can explore these
-related_words <- c("theory", "theories", "theoretical")
+library(tidyverse)
+library(quanteda)
+related_words <- c("phenomena", "phenomenon")
 map(related_words, ~colSums(read_in_the_abstracts_data$all_txts_c_dtm[, .x ]))
 #  so use the most common one, eg.:
 
@@ -30,6 +36,7 @@ map(related_words, ~colSums(read_in_the_abstracts_data$all_txts_c_dtm[, .x ]))
 # understanding
 # agency
 # practice
+# phenomenon
 
 # set the word here that we use in the later functions:
 target_feature <-  "theory"
@@ -59,7 +66,8 @@ insert_time_specific_token_data <-
 generate_embedding_matrix_data <- 
   generate_embedding_matrix_fn(insert_time_specific_token_data$all_txts_c_fcm,
                                insert_time_specific_token_data$time_specific_token,
-                               n_similar_words = 50)
+                               n_similar_words = 100,
+                               rank = 75)
 #-----------------------------------------------------------
 
 # Now we can draw the plot:
@@ -98,8 +106,11 @@ target_words <-
 ,"understanding"
 ,"agency"
 ,"practice"
+,"phenomenon"
   )
 
 
 map(target_words, 
-    ~insert_generate_plot_fn(.x, 5, read_in_the_abstracts_data))
+    ~insert_generate_plot_fn(.x, 
+                             year_interval= 10, 
+                             read_in_the_abstracts_data))

### 2020-05-22 documents/paper.Rmd
add text on temporal referencing and t-SNE

@@
 
-BM: Word embeddings have been used as evidence in studies of historical change in language and culture [@phillips2017intrinsic; @heuser2017word; @kulkarni2015statistically].
+BM: Word embeddings have been used as evidence in studies of historical change in language and culture [@phillips2017intrinsic; @heuser2017word; @kulkarni2015statistically; @pivovarova2019word].
 Semantic relations between words in a corpus are captured by the geometry of the vectors, so vectors can be compared with each other to find a word’s nearest neighbors [@giorcelli2019does].
@@
 @kulkarni2015statistically compute a linear transformation of words between any two time slices by solving a n-dimensional least squares problem of k nearest neighbor words (where n is the embedding dimension).
-@zhang2016past similarly use a linear transformation between a base and target time slices, and compute the
-linear transformation using anchor words that does not change meaning between the two time slices (requiringprior knowledge of words that are static).
+@zhang2016past similarly use a linear transformation between a base and target time slices, and compute the linear transformation using anchor words that does not change meaning between the two time slices (requiring prior knowledge of words that are static).
 @hamilton2016diachronic solves a n-dimensional Procrustes problem between every two adjacent time slices to perform the alignment.
+These alignment procedures are complex, and @dubossarsky2017outta show that they suffer from biases and produce spurious results.
+@dubossarsky2019time show that an alternative method, temporal referencing, avoids these concerns about previous alignment methods using a radically simpler procedure.
+Temporal referencing combines all time-specific corpora into one corpus, and generates word embeddings on this full corpus.
+Before generating the embeddings, they first replace each instance of the target word with a time-specific token.
+For example, the word 'theory' is replaced by the time-specfic token 'theory (1970-1980)' in the meeting abstracts from the years 1970-1980, and 'theory (1980-1990)' in the meeting abstracts from the years 1980-1990, and so on for each time slice.
+Then a single space is generated that contains a vector for each time-specific token, and the coordinates of the background, or context, words remain the same for each time slice, with no need for alignment.
+Besides the advantage of avoiding alignment, this temporal referencing method lowers data requirements (because context words are collapsed, and thus shared across time slices) and is suitable for smaller corpora, producing smooth change values without millions of words.
+A key assumption of this approach, as with the alignment methods, is that the semantics of the context words stay relatively stable over time.
 
-We then visualize the values across different dimensions to explore the historical trend of word meanings related to disciplinary change in archaeology.
+BM: Once the embeddings had been generated, we visualized the word vector values to explore the historical trends of word meanings in archaeology.
+We first subset the corpus to identify the words that are semantically similar to our target word using the cosine distance measure on pairs of word vectors.
+These similar words become the context words that map the semantic space that our target word moves through over time.
+Then we used the t-Distributed Stochastic Neighbor Embedding (t-SNE) method [@maaten2008visualizing] for constructing a two dimensional embedding of the word vectors to reduce the high dimensionality of the word vectors so they can projected on a scatter plot.
+t-SNE is a visualisation method that retaining the local structure of the data while also revealing some important global structure (such as clusters).
+This is useful for placing similar words close together on the plot, while maximizing the distance between dissimilar words.
+While t-SNE is highly flexible, and can often find structure where other dimensionality-reduction algorithms, such as PCA, cannot [@balamurali2016t], it can produce plots that are difficult to interpret, and some elements can be highly senstive to different hyperparameter settings, for example the size of the clusters and inter-cluster distances [@wattenberg2016how].
+A key limitation is that because t-SNE is non-linear (compared to PCA which is a linear transformation) and adapts to the underlying data, performing different transformations on different regions of the data, t-SNE plots are not easily interpretable in terms of the axes/units of the original, high-dimensional data.
 

### 2020-05-22 code/006-temp-ref-functions.R

@@ -53,6 +53,10 @@ interval <- floor_to_interval(as.numeric(year), year_interval)
 time_specific_token <- 
   paste0(target_feature, "_", interval, "_", interval + year_interval, "")
 
+# deal with the most recent one
+time_specific_token[length(time_specific_token)] <- 
+  str_replace(time_specific_token[length(time_specific_token)], "_\\d{4}$", "")
+
 all_txts_updated <- vector("character", length = length(all_txts))
 for(i in seq_len(length(all_txts))){
   
@@ -223,7 +227,10 @@ tsne_plot <-
             colour = "red",
             bg.color = "white", 
             bg.r = 0.15 ) +
-  theme_minimal()
+  theme_minimal() +
+  xlab("t-SNE dimension 1") +
+  ylab("t-SNE dimension 2") 
+  
 
 tsne_plot +
   ggtitle(paste0("Semantic shifts in the word '",

### 2020-05-22 code/006-temp-ref-scripts.R

@@ -40,7 +40,7 @@ map(related_words, ~colSums(read_in_the_abstracts_data$all_txts_c_dtm[, .x ]))
 
 # set the word here that we use in the later functions:
 target_feature <-  "theory"
-year_interval <-  5
+year_interval <-  7
 
 insert_generate_plot_fn <- 
   function(target_feature, 
@@ -67,7 +67,7 @@ generate_embedding_matrix_data <-
   generate_embedding_matrix_fn(insert_time_specific_token_data$all_txts_c_fcm,
                                insert_time_specific_token_data$time_specific_token,
                                n_similar_words = 100,
-                               rank = 75)
+                               rank = 30)
 #-----------------------------------------------------------
 
 # Now we can draw the plot:
@@ -88,7 +88,7 @@ temporal_referencing_plot_output <-
 
 ggsave(here::here(str_glue("figures/tr-plot-target-word-is-{target_feature}-{year_interval}-by-year-intervals.png")))
 
-  }
+}
 
 
 # Analyze and plot many target features in one action: 
@@ -107,10 +107,12 @@ target_words <-
 ,"agency"
 ,"practice"
 ,"phenomenon"
+,"evidence"
+,"data"
   )
 
 
 map(target_words, 
     ~insert_generate_plot_fn(.x, 
-                             year_interval= 10, 
+                             year_interval = 7, 
                              read_in_the_abstracts_data))

### 2020-05-27 documents/paper.Rmd
edit the background section

@@
 For example, @Garg2018 explored changes in gender stereotypes and attitudes toward ethnic minorities in the 20th and 21st centuries in the US.
-They created custom lists of target words representing gender and ethnicity, and neutral words, such as adjectives and jobs, serving as indicators for stereotypes and attitudes.
-<!--where did their word embeddings come from, and what algorithm did they use?
---> With word embeddings, they compared the association between each group <!--what is the group?
-we need to mention the groups before this sentence --> and neutral words, where the strength of association is calculated by computing average embedding distance.
-They found there is a job bias for both gender and ethnic minority, <!-- we really need a concrete example here --> this bias is significantly correlated with the actual job participation rates.
-In addition, there is a decreases in bias reflected by adjectives for ethnic minority and gender are associated with the women’s movement in the 1960s–1970s and Asian-American population growth in the 1960s and 1980s <!--"Another bias they identified was ...
-give the reader a concrete result from the paper  -->.
-This study demonstrates a successful application of word embeddings to detect social bias <!-- can we re-prhase to be about changing meanings of words?
---> and correlate the findings with historical events.
+They created word embeddings from different sources after 1900, including Google News, Google Books, and Corpus of Historical American English using Word2vex algorithm and New York Times, Wikipedia, and Common Crawl using GloVe algorithm.
+They made custom lists of group words representing gender and ethnicity, and neutral words, such as adjectives and jobs, serving as indicators for stereotypes and attitudes.
+For example, group words of man consists of he, man, boy, etc., and group words of Asian represented by last names such as li, wong, kim, etc.
+They compared the association between each group words (gender and ethnic minorities) and neutral words, where the strength of association is calculated by computing average embedding distance.
+Their results show that jobs such as nurses, dancers, housekeepers are biased toward women and there is an increase in adjectives for describing competence, such as intelligent, after women’s movement in the 1960s-1970s with higher participation rate in jobs.
+Another bias is the jobs of Asian-American were associated with academic positions reflecting the stereotype of the model minority, and their adjectives shift from negative words describing outsiders before 1950 to stereotype words about personality trait after 1950, with a growth of Asian-American population.
+This study demonstrates a successful use of word embeddings to detect semantic change in social bias over time with its sociocultural background.
 
-LW: Another example shows how word embeddings are effective <!-- Q: for what?
-A: for understanding semantic change over time --> by exploring the historical view on shared understandings of social classes in the 20th century in the US.
-@Kozlowski first compared the cultural dimensions created by word embeddings <!-- generated word embeddings from what data?
-and what algorithm did they use?
---> with a defined word list reflecting socio-economic differences, such as 'rich', 'poor' and so on.
-The high association <!-- of what?
-between what?
---> suggests that word embedding is a useful way to investigate people's views about social classes.
-They explored the relationship between dimensions of class identified in sociological theory, including morality, status, education, cultivation, gender, and employment.
-They found there is a high correlation <!-- of what?
-spell it out if this is about the word vectors, say --> between dimensions of education, cultivation, and morality, which corresponds to the owner/worker relation <!-- what relation?
-spell out more --> defined by classic Marxism.
-In addition, they found a stable relationship between different cultural dimensions of class over the century even though there was a major shift in the condition of economy, industry, and employment in the present day.
-This case study indicates the robustness of word embeddings as a method to examine historical trends of social classes.
+LW: Another example shows how word embeddings are effective for identifying semantic changes in shared understandings of social classes in the 20th century in the US.
+@Kozlowski analyzed word embeddings generated on millions of books from Google Ngrams texts between 1990 and 2012 using the word2Vec algorithm with the GloVe algorithm as a cross-validation.
+The results suggest that 'employment', one of the class dimensions, shows most of the semantic changes from titles of formal office, such as 'lords' and 'governor', to 'promoter' and 'mogul` referring to economic status with power.
+Similarly, the class dimension 'employee' shifts from 'wage' and 'earners' in the early 20th century, 'soldier' during World War I, to 'retirement' and 'qualified' in the end of the century, reflecting the emergence of middle-class.
+They also found a close relationship between word embeddings in the classes, including education, morality, and cultivation, over time despite there is a major shift in the condition of economy, industry, and employment in the present day.
+This case study indicates the robustness of word embeddings as a method to examine historical trends by exploring social classes.
 

### 2020-05-28 documents/paper.Rmd
update method section

@@
 LW: We generated the word embeddings on the co-occurrence matrix using the GloVe algorithm from the text2vec package for R [@Pennington2014; @Selivanov2016].
-GloVe (Global Vectors) is a unsupervised learning model that creates word vectors to represent semantic regularity and capture relationships between words [@Pennington2014; @Garg2018].
+GloVe (Global Vectors) is an unsupervised learning model that creates word vectors to represent semantic regularity and capture relationships between words [@Pennington2014; @Garg2018].
+The general idea underlying the model is that the multiplied value of word vectors i and j will meet the logarithm of the co-occurrence count of word i and j in the matrix.
+Following this principle, GloVe performs a logistic regression continuously for every word along the text to decide vector values by maximizing the probability of a surrounding word with the occurrence of a center word.
+It involves three steps to complete this task.
+First, it adds intercepts associated with each word in a pair to compensate for the difference between the average of a word vector and its weighted sum.
+Second, it minimizes the sum of all squared errors from the previous step by giving weight to vectors with a function.
+Third, it deals with word pairs with large co-occurrence count to avoid a skewed number.
+Finally, GloVe generates two sets of word vectors as the roles of words (i.e.
+as a center or a surrounding word) in each pair can be reversed when GloVe algorithm moving through.
+The final output of word vectors is the sum of these two sets of values.
 

### 2020-05-28 documents/paper.Rmd
notes from class

--- 2020-05-27 documents/paper.Rmd
@@
 
-LW:  Word embeddings have also been used to track demographic and occupational social shifts and gender stereotypes.
+LW:  Word embeddings have also been used to track demographic and occupational shifts and gender stereotypes.
 For example, @Garg2018 explored changes in gender stereotypes and attitudes toward ethnic minorities in the 20th and 21st centuries in the US.
-They created word embeddings from different sources after 1900, including Google News, Google Books, and Corpus of Historical American English using Word2vex algorithm and New York Times, Wikipedia, and Common Crawl using GloVe algorithm.
-They made custom lists of group words representing gender and ethnicity, and neutral words, such as adjectives and jobs, serving as indicators for stereotypes and attitudes.
-For example, group words of man consists of he, man, boy, etc., and group words of Asian represented by last names such as li, wong, kim, etc.
-They compared the association between each group words (gender and ethnic minorities) and neutral words, where the strength of association is calculated by computing average embedding distance.
-Their results show that jobs such as nurses, dancers, housekeepers are biased toward women and there is an increase in adjectives for describing competence, such as intelligent, after women’s movement in the 1960s-1970s with higher participation rate in jobs.
-Another bias is the jobs of Asian-American were associated with academic positions reflecting the stereotype of the model minority, and their adjectives shift from negative words describing outsiders before 1950 to stereotype words about personality trait after 1950, with a growth of Asian-American population.
-This study demonstrates a successful use of word embeddings to detect semantic change in social bias over time with its sociocultural background.
+They used word embeddings from different sources after 1900, including Google News, Google Books, and the Corpus of Historical American English using the word2vec algorithm and New York Times using the GloVe algorithm.
+They extracted word embeddings for sets of words representing gender (e.g.
+he, man, boy, etc.) and ethnicity (e.g.
+last names such as Li, Wong, Kim), and a set of neutral words, such as adjectives and job titles, serving as indicators for stereotypes and attitudes.
+They compared the association of word vectors between each set words (gender and ethnic minorities) and neutral words, where the strength of association is calculated by computing average embedding distance.
+Their results show that jobs such as nurses, dancers, housekeepers have been biased toward women and there is an increase in adjectives for describing women's competence, such as intelligent, after the women’s movement in the 1960s-1970s, with higher participation rates in jobs.
+Another bias is the jobs of Asian-Americans which were associated with academic positions, reflecting the stereotype of the model minority, and their adjectives shift from negative words describing outsiders before 1950 to stereotype words about personality traits after 1950, with the growth of the Asian-American population.
+This study demonstrates a successful use of word embeddings to detect semantic change over time that reveal social biases that correspond to broader sociocultural contexts.
 
-LW: Another example shows how word embeddings are effective for identifying semantic changes in shared understandings of social classes in the 20th century in the US.
-@Kozlowski analyzed word embeddings generated on millions of books from Google Ngrams texts between 1990 and 2012 using the word2Vec algorithm with the GloVe algorithm as a cross-validation.
-The results suggest that 'employment', one of the class dimensions, shows most of the semantic changes from titles of formal office, such as 'lords' and 'governor', to 'promoter' and 'mogul` referring to economic status with power.
-Similarly, the class dimension 'employee' shifts from 'wage' and 'earners' in the early 20th century, 'soldier' during World War I, to 'retirement' and 'qualified' in the end of the century, reflecting the emergence of middle-class.
+LW: Another example shows how word embeddings are effective for identifying semantic changes in shared understandings of US social classes in the 20th century.
+@Kozlowski used the word2vec algorithm to generate word embeddings from millions of books published between between 1900 and 2012 that are available in the Google Ngrams corpus.
+<!-- need to prepare the reader: what are the class dimension and how were they identified?
+capture a sense of that they project word embeddings on a project for each cultural dimension --> The results suggest that 'employment', one of the class dimensions, shows substantial semantic changes.
+For example, titles of formal office, such as 'lords' and 'governor', change to 'promoter' and 'mogul', which are positions of economic status with power.
+Similarly, the class dimension 'employee' shifts from an association with words like 'wage' and 'earners' in the early 20th century, and 'soldier' during World War I, to 'retirement' and 'qualified' in the end of the century.
+This reflects the emergence of a middle-class in the US.
 They also found a close relationship between word embeddings in the classes, including education, morality, and cultivation, over time despite there is a major shift in the condition of economy, industry, and employment in the present day.
 This case study indicates the robustness of word embeddings as a method to examine historical trends by exploring social classes.
+<!-- check all the sentences to make sure the main ideas flow nicely through the entire paragraph -->
 
@@
 
-LW: We generated the word embeddings on the co-occurrence matrix using the GloVe algorithm from the text2vec package for R [@Pennington2014; @Selivanov2016].
-GloVe (Global Vectors) is a unsupervised learning model that creates word vectors to represent semantic regularity and capture relationships between words [@Pennington2014; @Garg2018].
-
-<!-- one paragraph of plain and simple description of how Glove computes word vectors from co-oc matrix -->
+LW: We generated the word embedding on the co-occurrence matrix using the GloVe algorithm from the text2vec package for R [@Pennington2014; @Selivanov2016].
+GloVe (Global Vectors) is an unsupervised learning algorithm that creates word vectors to represent semantic regularities and capture relationships between words [@Pennington2014].
+The general idea is that the multiplied value of the final word vectors for any two words will be equivalent to the logarithm of those words' probability of co-occurrence.
+To achieve this, GloVe performs a weighted least squares regression for every word continuously along the text to generate vector values by maximizing the probability of a surrounding word with the occurrence of a center word.
+GloVe generates two sets of word vectors for each word (i.e.
+for the word as a center word or a surrounding word), and final word vector is the sum of these two sets of values.
 
@@
 Before generating the embeddings, they first replace each instance of the target word with a time-specific token.
-For example, the word 'theory' is replaced by the time-specfic token 'theory (1970-1980)' in the meeting abstracts from the years 1970-1980, and 'theory (1980-1990)' in the meeting abstracts from the years 1980-1990, and so on for each time slice.
+For example, the word 'theory' is replaced by the time-specific token 'theory (1970-1980)' in the meeting abstracts from the years 1970-1980, and 'theory (1980-1990)' in the meeting abstracts from the years 1980-1990, and so on for each time slice.
 Then a single space is generated that contains a vector for each time-specific token, and the coordinates of the background, or context, words remain the same for each time slice, with no need for alignment.
@@
 - basic details (e.g. word and abstract counts)
-- word frequency time series
+-- word frequency time series
 - TR plots

### 2020-05-28 code/006-temp-ref-functions.R

@@ -35,7 +35,8 @@ return(list(all_txts = all_txts,
 
 insert_time_specific_token_fn <- function(all_txts = all_txts,
                                           target_feature = "mechanisms",
-                                          year_interval = 5){
+                                          year_interval = 5,
+                                          min_termfreq = 5){
 
 # https://www.aclweb.org/anthology/P19-1044.pdf is our inspiration 
 # in each text we want to replace the target feature (word) with a tagged feature,
@@ -53,9 +54,10 @@ interval <- floor_to_interval(as.numeric(year), year_interval)
 time_specific_token <- 
   paste0(target_feature, "_", interval, "_", interval + year_interval, "")
 
-# deal with the most recent one
+# deal with the most recent one by grouping it with the previous one
+# otherwise we get 2020 all by itself
 time_specific_token[length(time_specific_token)] <- 
-  str_replace(time_specific_token[length(time_specific_token)], "_\\d{4}$", "")
+  time_specific_token[(length(time_specific_token) - 1)]
 
 all_txts_updated <- vector("character", length = length(all_txts))
 for(i in seq_len(length(all_txts))){
@@ -80,7 +82,7 @@ all_txts_c_dtm_toks1 <- tokens(all_txts_updated_c,
                   remove_numbers = TRUE,
                   remove_url = FALSE,
                   remove_separators = TRUE,
-                  split_hyphens = FALSE)
+                  split_hyphens = FALSE) 
 
 all_txts_c_feats <- dfm(all_txts_c_dtm_toks1, 
                 tolower = TRUE,
@@ -88,7 +90,7 @@ all_txts_c_feats <- dfm(all_txts_c_dtm_toks1,
                # remove = c(stopwords("english")),
                # stem = TRUE, 
                 remove_punct = TRUE) %>%
-  dfm_trim(min_termfreq = 3) %>%
+  dfm_trim(min_termfreq = min_termfreq) %>% # try to limit the exotic words that are only found in earlier years
   featnames()
 
 # leave the pads so that non-adjacent words will not become adjacent
@@ -209,14 +211,14 @@ tsne_plot <-
             aes(x = V1, 
                 y = V2, 
                 label = word), 
-            size = 2,
-            check_overlap = TRUE) +
+            size = 1.5,
+            check_overlap = FALSE) +
   geom_segment(data = tsne_plot_target_features, 
                aes(x = V1, 
[…]

### 2020-05-28 code/006-temp-ref-scripts.R

@@ -38,14 +38,11 @@ map(related_words, ~colSums(read_in_the_abstracts_data$all_txts_c_dtm[, .x ]))
 # practice
 # phenomenon
 
-# set the word here that we use in the later functions:
-target_feature <-  "theory"
-year_interval <-  7
-
 insert_generate_plot_fn <- 
-  function(target_feature, 
-           year_interval, 
-           read_in_the_abstracts_data){
+  function(target_feature = "theory", 
+           year_interval = 5, 
+           read_in_the_abstracts_data,
+           min_termfreq = 15){
 
 # Now we insert the time-specific token, like "theory_1975_1980"
 
@@ -55,7 +52,8 @@ insert_generate_plot_fn <-
 insert_time_specific_token_data <- 
   insert_time_specific_token_fn(read_in_the_abstracts_data$all_txts,
                                 target_feature,
-                                year_interval)
+                                year_interval,
+                                min_termfreq)
 #-----------------------------------------------------------
 
 # Now we can generate the word embeddings for our target feature:
@@ -81,7 +79,7 @@ temporal_referencing_plot_output <-
                                time_specific_token = insert_time_specific_token_data$time_specific_token,
                                target_feature,
                                year_interval,
-                               perplexity = 50)
+                               perplexity = 100)
 #-----------------------------------------------------------
 
 # Now we can save the plot:
@@ -95,24 +93,26 @@ ggsave(here::here(str_glue("figures/tr-plot-target-word-is-{target_feature}-{yea
 
 target_words <- 
   c(
- "model"
-,"theory"
-,"mechanisms"
-,"hypothesis"
-,"explanations"
-,"inference"
-,"narrative"
-,"interpretation"
-,"understanding"
-,"agency"
-,"practice"
-,"phenomenon"
-,"evidence"
-,"data"
+ "theory"
+# ,"model"
+# ,"mechanisms"
+# ,"hypothesis"
+# ,"explanations"
+# ,"inference"
+# ,"narrative"
+# ,"interpretation"
+# ,"understanding"
+# ,"agency"
+# ,"practice"
+# ,"phenomenon"
+# ,"evidence"
+# ,"data"
   )
 
+library(tidyverse)
+library(quanteda)
 
 map(target_words, 
     ~insert_generate_plot_fn(.x, 
-                             year_interval = 7, 
+                             year_interval = 5, 
                              read_in_the_abstracts_data))

### 2020-05-29 documents/paper.Rmd
edit the para in the background section

@@
 LW: Another example shows how word embeddings are effective for identifying semantic changes in shared understandings of US social classes in the 20th century.
-@Kozlowski used the word2vec algorithm to generate word embeddings from millions of books published between between 1900 and 2012 that are available in the Google Ngrams corpus.
-<!-- need to prepare the reader: what are the class dimension and how were they identified?
-capture a sense of that they project word embeddings on a project for each cultural dimension --> The results suggest that 'employment', one of the class dimensions, shows substantial semantic changes.
+@Kozlowski used the word2vec algorithm to generate word embeddings from millions of books published between 1900 and 2012 that are available in the Google Ngrams corpus.
+They first constructed cultural dimensions of class that commonly discussed in sociological theory from word embeddings, including affluence, occupation, education, states, and cultivated taste.
+They then projected word embeddings that consist of sets of antonym pairs on a project for each cultural dimension as new dimensions to see their relationships.
+For instance, dimensions of rich-poor, priceless-worthless, and prosperous-bankrupt were projected on to the affluence dimension.
+The most interesting result is that 'employment', the occupation dimension, shows substantial semantic changes.
 For example, titles of formal office, such as 'lords' and 'governor', change to 'promoter' and 'mogul', which are positions of economic status with power.
-Similarly, the class dimension 'employee' shifts from an association with words like 'wage' and 'earners' in the early 20th century, and 'soldier' during World War I, to 'retirement' and 'qualified' in the end of the century.
+Similarly, 'employee', another occupation dimension, shifts from an association with words like 'wage' and 'earners' in the early 20th century, and 'soldier' during World War I, to 'retirement' and 'qualified' in the end of the century.
 This reflects the emergence of a middle-class in the US.
-They also found a close relationship between word embeddings in the classes, including education, morality, and cultivation, over time despite there is a major shift in the condition of economy, industry, and employment in the present day.
+They also found a close relationship between word pairs across different cultural dimensions, including education, morality, and cultivated taste, over time despite there is a major shift in the condition of economy, industry, and employment in the present day.
 This case study indicates the robustness of word embeddings as a method to examine historical trends by exploring social classes.
-<!-- check all the sentences to make sure the main ideas flow nicely through the entire paragraph -->
 

### 2020-06-04 documents/paper.Rmd
work on basic details in result section

@@
 saa_data <- readxl::read_excel(here::here("data", "raw-data", "saa-abstracts-tally.xlsx"))
+```
 
+```{r basic-stats}
 # basic stats
@@
 
+# total word counts
 all_word_counts <- sum(stringi::stri_count_words(all_txts))
+
+# the hightst number of abstract
+high_abs <- max(saa_data$number_of_abstracts, na.rm = TRUE)
 ```
@@
 - basic details (e.g. word and abstract counts)
+
+LW: Figure XX shows that there is an increase in the abstract counts over time that can be divided into two groups, a group of an early time from 1962 to 1994 with abstracts less than 1,000 per year, and a group of a later time from 2004 to 2020 with abstracts more than 1,500 per year.
+We observed the highest abstract counts (`r prettyNum(high_abs, big.mark = ",")`) in 2015, followed by 2017, 2019, 2018, and 2014 with more than 4,000 attracts, but a drop in 2020 with around 3,000 abstracts.
+The total number of word counts per year shows a similar pattern of growth over time that corresponds to the trend of the abstract counts.
+This suggests that the more abstract counts, the more word counts per year that matches our expectation for the data.
+
+LW: We compared the word frequencies between two time periods, before 2004 and after 2004, to explore the general change of the usage of words in archaeology over time.
+We observed (Figure XX) an interesting pattern that shows a considerable high proportion of words used less often compared to words used more often after 2004, despite the word count is much higher in the later time.
+That is, archaeologists in more recent years have a smaller range of word choices and tend to not use some words, such as prehistoric, data, site, settlement, and patterns, used frequently in the past.
+In contrast, the words such as landscape, ancient, social, practices, and understanding become common after 2004.
+It could reflect an extinction of a large group of words during the course of the development of archaeology.
+One reason may be researchers have a more common idea about the selection of words when the community becomes bigger and more international.
+This leads to the use of a smaller group of words that present the sharing understanding about archaeology instead of the use of large group of obscure words demonstrate one's knowledge.
+
 -- word frequency time series

### 2020-06-04 documents/paper.Rmd
fixed typos

@@
 This reflects the emergence of a middle-class in the US.
-They also found a close relationship between word pairs across different cultural dimensions, including education, morality, and cultivated taste, over time despite there is a major shift in the condition of economy, industry, and employment in the present day.
-This case study indicates the robustness of word embeddings as a method to examine historical trends by exploring social classes.
+They also found a stable, close relationship between word pairs across different cultural dimensions, including education, morality, and cultivated taste, over time despite there being a major shift in the condition of economy, industry, and employment in the present day.
+This case study indicates word embeddings is a robust method for examining historical trends of semantic meaning for social class.
 
@@
 BM: One of the key practical challenges of studying temporal change in word embeddings is aligning the vectors representing the same words across multiple time periods.
-This is because the stochastic nature of word embeddings algorithms makes them generate representations that always have different coordinate systems and thus one set of embeddings are are not directly comparable to another set generated on a different corpus (i.e.
+This is because the stochastic nature of word embeddings algorithms makes them generate representations that always have different coordinate systems and thus one set of embeddings are not directly comparable to another set generated on a different corpus (i.e.
 from a different time period).
@@
 These similar words become the context words that map the semantic space that our target word moves through over time.
-Then we used the t-Distributed Stochastic Neighbor Embedding (t-SNE) method [@maaten2008visualizing] for constructing a two dimensional embedding of the word vectors to reduce the high dimensionality of the word vectors so they can projected on a scatter plot.
+Then we used the t-Distributed Stochastic Neighbor Embedding (t-SNE) method [@maaten2008visualizing] for constructing a two dimensional embedding of the word vectors to reduce the high dimensionality of the word vectors so they can be projected on a scatter plot.
 t-SNE is a visualisation method that retaining the local structure of the data while also revealing some important global structure (such as clusters).
 This is useful for placing similar words close together on the plot, while maximizing the distance between dissimilar words.
-While t-SNE is highly flexible, and can often find structure where other dimensionality-reduction algorithms, such as PCA, cannot [@balamurali2016t], it can produce plots that are difficult to interpret, and some elements can be highly senstive to different hyperparameter settings, for example the size of the clusters and inter-cluster distances [@wattenberg2016how].
+While t-SNE is highly flexible, and can often find structure where other dimensionality-reduction algorithms, such as PCA, cannot [@balamurali2016t], it can produce plots that are difficult to interpret, and some elements can be highly sensitive to different hyperparameter settings, for example the size of the clusters and inter-cluster distances [@wattenberg2016how].
 A key limitation is that because t-SNE is non-linear (compared to PCA which is a linear transformation) and adapts to the underlying data, performing different transformations on different regions of the data, t-SNE plots are not easily interpretable in terms of the axes/units of the original, high-dimensional data.
@@
 LW: We compared the word frequencies between two time periods, before 2004 and after 2004, to explore the general change of the usage of words in archaeology over time.
-We observed (Figure XX) an interesting pattern that shows a considerable high proportion of words used less often compared to words used more often after 2004, despite the word count is much higher in the later time.
+We observed (Figure XX) an interesting pattern that shows a considerably high proportion of words used less often compared to words used more often after 2004, despite the word count being much higher in the later time.
 That is, archaeologists in more recent years have a smaller range of word choices and tend to not use some words, such as prehistoric, data, site, settlement, and patterns, used frequently in the past.
-In contrast, the words such as landscape, ancient, social, practices, and understanding become common after 2004.
+In contrast, words such as landscape, ancient, social, practices, and understanding became common after 2004.
 It could reflect an extinction of a large group of words during the course of the development of archaeology.
 One reason may be researchers have a more common idea about the selection of words when the community becomes bigger and more international.
-This leads to the use of a smaller group of words that present the sharing understanding about archaeology instead of the use of large group of obscure words demonstrate one's knowledge.
+This leads to the use of a smaller group of words that present the sharing understanding about archaeology instead of the use of a large group of obscure words that demonstrate one's knowledge.
 

### 2021-07-19 documents/paper.Rmd
update

@@
 In the 1980s an external, critical history phase began, seeking to understand the interactions between archaeological practice and its political, economic and social contexts.
+<!-- need to expand more on 'internalist phase' and 'critical history' phase...
+define and give examples...
+-->
 
-BM: Among historians of archaeology there are two major phases in  ‘critical history of archaeology’ in the 1980s had been characterized by the replacement of an
-
-- previous efforts at the history of archaeology (e.g. Bruce Trigger's book)
-- previous efforts at text analysis in archaeology (e.g. bibliometrics, citation networks)
-- new phase: we are now in the third phase  - computational text analysis
-- one promising new method is WE... successful applications of WE in other fields...
-
-- background to Word embeddings
+BM: We believe there is now a third phase in studying the history of archaeology...
+quantiative analysis of text...
+previous efforts at text analysis in archaeology (e.g.
+bibliometrics, citation networks) <!--  define and give examples...
+we got many of these listed in our notes already -->
 
@@
 
-LW: Another example shows how word embeddings are effective for identifying semantic changes in shared understandings of US social classes in the 20th century.
-@Kozlowski used the word2vec algorithm to generate word embeddings from millions of books published between 1900 and 2012 that are available in the Google Ngrams corpus.
-They first constructed cultural dimensions of class that commonly discussed in sociological theory from word embeddings, including affluence, occupation, education, states, and cultivated taste.
-They then projected word embeddings that consist of sets of antonym pairs on a project for each cultural dimension as new dimensions to see their relationships.
-For instance, dimensions of rich-poor, priceless-worthless, and prosperous-bankrupt were projected on to the affluence dimension.
-The most interesting result is that 'employment', the occupation dimension, shows substantial semantic changes.
+LW: @Kozlowski shows how word embeddings are effective for identifying semantic changes in shared understandings of US social classes in the 20th century.
+They used the word2vec algorithm to generate word embeddings from millions of books published between 1900 and 2012 that are available in the Google Ngrams corpus.
+The constructed cultural dimensions by word vector arithmetic operations on antonym pairs relating to key cultural concepts.
+They computed an 'affluence dimension' using arithmetic operations on word vectors, e.g.
+'rich-poor', 'priceless-worthless', and 'prosperous-bankrupt', etc..
+The resulting values capture a semantic sense of affluence and were used to draw a line on the plot that is the affluence dimension that other word vectors are projected on.
+A major finding is that employment words show substantial semantic changes on the affluence dimension over time.
 For example, titles of formal office, such as 'lords' and 'governor', change to 'promoter' and 'mogul', which are positions of economic status with power.
-Similarly, 'employee', another occupation dimension, shifts from an association with words like 'wage' and 'earners' in the early 20th century, and 'soldier' during World War I, to 'retirement' and 'qualified' in the end of the century.
+Similarly, employee words shift from an association with words like 'wage' and 'earners' in the early 20th century, and 'soldier' during World War I, to 'retirement' and 'qualified' in the end of the century.
 This reflects the emergence of a middle-class in the US.
-They also found a stable, close relationship between word pairs across different cultural dimensions, including education, morality, and cultivated taste, over time despite there being a major shift in the condition of economy, industry, and employment in the present day.
-This case study indicates word embeddings is a robust method for examining historical trends of semantic meaning for social class.
+This case study indicates word embeddings can be a robust method for examining historical trends in semantic change for words relating to social class.
 
@@
 
-# the hightst number of abstract
+# the highest number of abstract
 high_abs <- max(saa_data$number_of_abstracts, na.rm = TRUE)
@@
 BM: One of the key practical challenges of studying temporal change in word embeddings is aligning the vectors representing the same words across multiple time periods.
-This is because the stochastic nature of word embeddings algorithms makes them generate representations that always have different coordinate systems and thus one set of embeddings are not directly comparable to another set generated on a different corpus (i.e.
+This is because the stochastic nature of word embeddings algorithms makes them generate representations that always have different coordinate systems.
+Embeddings generated from one corpus are not directly comparable to embeddings on the same words but generated on a different corpus (i.e.
 from a different time period).
@@
 For example, the word 'theory' is replaced by the time-specific token 'theory (1970-1980)' in the meeting abstracts from the years 1970-1980, and 'theory (1980-1990)' in the meeting abstracts from the years 1980-1990, and so on for each time slice.
-Then a single space is generated that contains a vector for each time-specific token, and the coordinates of the background, or context, words remain the same for each time slice, with no need for alignment.
+Then a single coordinate system is generated that contains a vector for each time-specific token, and the coordinates of the background, or context, words remain the same for each time slice, with no need for alignment.
 Besides the advantage of avoiding alignment, this temporal referencing method lowers data requirements (because context words are collapsed, and thus shared across time slices) and is suitable for smaller corpora, producing smooth change values without millions of words.
@@
 
-BM: Once the embeddings had been generated, we visualized the word vector values to explore the historical trends of word meanings in archaeology.
-We first subset the corpus to identify the words that are semantically similar to our target word using the cosine distance measure on pairs of word vectors.
+BM: Once the embeddings have been generated, we visualized the word vector values to explore the historical trends of word meanings in archaeology.
+We identified words that are semantically similar to our target word by computing the cosine distance measure on all pairs of word vectors, keeping the closest X words.
 These similar words become the context words that map the semantic space that our target word moves through over time.
@@
 This is useful for placing similar words close together on the plot, while maximizing the distance between dissimilar words.
+A key limitation is that because t-SNE is non-linear (compared to PCA which is a linear transformation) and adapts to the underlying data, performing different transformations on different sub-groups of the data, t-SNE plots are not easily interpretable in terms of the axes/units of the original, high-dimensional data.
 While t-SNE is highly flexible, and can often find structure where other dimensionality-reduction algorithms, such as PCA, cannot [@balamurali2016t], it can produce plots that are difficult to interpret, and some elements can be highly sensitive to different hyperparameter settings, for example the size of the clusters and inter-cluster distances [@wattenberg2016how].
-A key limitation is that because t-SNE is non-linear (compared to PCA which is a linear transformation) and adapts to the underlying data, performing different transformations on different regions of the data, t-SNE plots are not easily interpretable in terms of the axes/units of the original, high-dimensional data.
 
@@
 
-```{r temporal-referencing-plots, cache.extra=tools::md5sum(here::here("006-testing-word-2-vec-temporal-referencing-by-period.R"))}
-source(here::here("code/006-testing-word-2-vec-temporal-referencing-by-period.R"))
+# Results
+
+```{r time-series, cache.extra=tools::md5sum(here::here("code/004-basic-word-stats-and-plots.R"))}
+# show plot of abstract and word count
+source(here::here("code/004-basic-word-stats-and-plots.R"))
+knitr::include_graphics(here::here("figures/time-series-count-abstracts-and-words.png"))
+```
+
+```{r basic-word-change, cache.extra=tools::md5sum(here::here("code/005-basic-word-stats-and-plots.R"))}
+# show plot of basic word change
+source(here::here("code/005-basic-word-stats-and-plots.R"))
+knitr::include_graphics(here::here("figures/time-series-count-abstracts-and-words.png"))
+```
+
+```{r temporal-referencing-fns, cache.extra=tools::md5sum(here::here("code/006-temp-ref-functions.R"))}
+# load the functions for computing temporal referencing and drawing plots
+source(here::here("code/006-temp-ref-functions.R"))
+```
+
+```{r temporal-referencing-plots, cache.extra=tools::md5sum(here::here("code/006-006-temp-ref-scripts.R"))}
+# run the functions for computing temporal referencing and drawing plots
+source(here::here("code/006-temp-ref-scripts.R"))
+```
+
+```{r temporal-referencing-plot-theory, cache.extra=tools::md5sum(here::here("figures/tr-plot-target-word-is-theory-5-by-year-intervals.png"))}
+knitr::include_graphics(here::here("figures/tr-plot-target-word-is-theory-5-by-year-intervals.png"))
 ```
@@
 LW: Figure XX shows that there is an increase in the abstract counts over time that can be divided into two groups, a group of an early time from 1962 to 1994 with abstracts less than 1,000 per year, and a group of a later time from 2004 to 2020 with abstracts more than 1,500 per year.
-We observed the highest abstract counts (`r prettyNum(high_abs, big.mark = ",")`) in 2015, followed by 2017, 2019, 2018, and 2014 with more than 4,000 attracts, but a drop in 2020 with around 3,000 abstracts.
+We observed the highest abstract counts (n = `r prettyNum(high_abs, big.mark = ",")`) in 2015, followed by 2017, 2019, 2018, and 2014 with more than 4,000 attracts, but a drop in 2020 with around 3,000 abstracts.
 The total number of word counts per year shows a similar pattern of growth over time that corresponds to the trend of the abstract counts.
 This suggests that the more abstract counts, the more word counts per year that matches our expectation for the data.
+<!-- more interesting to say "This suggest that the number of words per abstract has been very stable over time for the SAA meetings -->
 
-LW: We compared the word frequencies between two time periods, before 2004 and after 2004, to explore the general change of the usage of words in archaeology over time.
-We observed (Figure XX) an interesting pattern that shows a considerably high proportion of words used less often compared to words used more often after 2004, despite the word count being much higher in the later time.
-That is, archaeologists in more recent years have a smaller range of word choices and tend to not use some words, such as prehistoric, data, site, settlement, and patterns, used frequently in the past.
-In contrast, words such as landscape, ancient, social, practices, and understanding became common after 2004.
-It could reflect an extinction of a large group of words during the course of the development of archaeology.
+LW: We compared the word frequencies between two time periods, before 2004 and after 2004, to explore changes in the use of words in archaeology over time.
+We observed (Figure XX) an interesting pattern that shows a considerably higher proportion of words used less often, compared to words used more often after 2004, despite the word count being much higher in the later time.
+That is, archaeologists in more recent years have a smaller vocabulary and tend to not use some words, such as 'prehistoric', 'data', 'site', 'settlement', and 'patterns', as frequently as they did in the past.
+In contrast, words such as 'landscape', 'ancient', 'social', 'practices', and 'understanding' became common after 2004.
+It could reflect an extinction of a large group of words during the development of archaeology.
 One reason may be researchers have a more common idea about the selection of words when the community becomes bigger and more international.
-This leads to the use of a smaller group of words that present the sharing understanding about archaeology instead of the use of a large group of obscure words that demonstrate one's knowledge.
+This leads to the use of a smaller group of words that present the shared understanding about archaeology instead of the use of a large group of obscure words that demonstrate one's knowledge.
+<!-- sounds like a value judgement -->
 

### 2021-07-19 code/004-basic-word-stats-and-plots.R

@@ -1,105 +1,4 @@
-# Words per year
 
-library(tidyverse)
-
-# starting with https://quanteda.io/articles/quickstart.html
-library("quanteda")
-
-# get all txt files of abstracts
-abstract_txt_files <- 
-  fs::dir_ls(here::here("data/derived-data/"), 
-             recurse = TRUE, 
-             regexp = '\\.txt$')
-
-all_txts <- tolower(readtext::readtext(abstract_txt_files))
-names(all_txts) <- map_chr(names(all_txts), ~str_match_all(.x, "\\d{4}")[[1]][1,1])
-
-# word count checking
-all_txts_word_count_per_year <- 
-map_int(all_txts, 
-       stringi::stri_count_words) %>% 
-  stack %>% 
-  as_tibble() %>% 
-  rename(words = values,
-         year = ind) %>% 
-  mutate(year = as.numeric(as.character(year)))
-
-# count all words for each year
-all_txts_c <- corpus(all_txts)
-all_txts_c_summary <- 
-  summary(all_txts_c) %>%   
-  mutate(year = parse_number(Text))
-  
-# make a dfm
-all_txts_c_dtm <- 
-  dfm(all_txts_c,
-      remove_punct = TRUE)
-
-# plot of all words in each year
-plot_all_words <- 
-  ggplot(data = all_txts_c_summary, 
-         aes(x = year , 
-             y = Tokens / 1000, 
-             group = 1)) +
-  geom_col() +
-  scale_x_continuous(labels = c(seq(1940, 2020, 2)), 
-                     breaks = seq(1940, 2020, 2),
-                     name = "Year") +
-  scale_y_continuous(name = "Total word count per year (x 1,000)",
-                     labels = scales::comma) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, 
-                                   vjust = 0.5)) 
-  
-# Our key words over time
-  
-keywords <- 
-  c("identity", 
-    "power", 
-    "awareness", 
-    "transformation", 
-    "presentation", 
-    "cognition")
-
-dfm_keywords <- 
-  dfm_select(all_txts_c_dtm, 
-             pattern = keywords, 
-             selection = "keep")
-
-dfm_keywords_tbl <- 
-  convert(dfm_keywords, to = "data.frame") %>% 
-  pivot_longer(-document, 
-               names_to = "keyword",
-               values_to = "n") %>% 
[…]

### 2021-07-19 code/0041-basic-word-stats-and-plots.R

@@ -1,145 +1,0 @@
-df <- all_txts_c_dtm
-
-dont_care <- 
-  c('university', 'abstracts', 'meeting', 
-    'annual', 'paper', 'chair', 'across', 
-    'presented', 'session', 'college', 'area',
-    'discussed', 'this', 'that', 'there', 'which',
-    'their', 'poster', 'using', 'through', 'into',
-    'some', 'from', 'been', 'discuss', 'while',
-    'over')
-
-# from tidytext
-df_tbl <- 
-  convert(df, to = "data.frame") %>% 
-  pivot_longer(-document, 
-               names_to = "word",
-               values_to = "n") %>% 
-  mutate(year = parse_number(document)) %>% 
-  filter(word %in% names(data_int_syllables)) %>% 
-  filter(str_length(word) > 3) %>% 
-  filter(!word %in% dont_care)
-
-df_tbl_yearly_totals <- 
-  df_tbl %>% 
-  group_by(year) %>% 
-  summarise(total_words = sum(n, na.rm = TRUE))
-  
-  
-# compute proportion of all words per year
-df_tbl_prop <- 
-  df_tbl %>% 
- # left_join(all_txts_c_summary) %>% 
-  left_join(df_tbl_yearly_totals) %>% 
-  mutate(prop = n / total_words ) %>% 
-  filter(prop > 0) %>% 
-  group_by(word) %>% 
-  mutate(sum_the_word = sum(n)) %>% 
-  mutate(keyword_n = str_c(word, " (n = ", sum_the_word, ")"))
-
-# want to find words with the greatest difference in 
-# proportion between two time periods 
-
-df_tbl_prop_two_groups <- 
-  df_tbl_prop %>% 
-  mutate(year_group = ifelse(document < 2004, 
-                             "early", 
-                             "late"))
-
-df_tbl_prop_two_groups_diff <- 
-df_tbl_prop_two_groups %>% 
-  group_by(word, year_group) %>% 
-  summarise(mean_prop = mean(prop, na.rm = TRUE)) %>% 
-  pivot_wider(names_from = year_group,
-              values_from = mean_prop) %>% 
-  drop_na() %>% 
-  mutate(diff = late - early)
-
-df_tbl_prop_two_groups_diff_interesting <- 
-df_tbl_prop_two_groups_diff %>% 
-  filter(diff >= 0.00001 | diff <= -0.00001)
-
-df_tbl_prop_two_groups_diff_interesting_top <- 
-  df_tbl_prop_two_groups_diff_interesting %>% 
[…]

### 2021-07-19 code/005-basic-word-stats-and-plots.R

@@ -1,0 +1,148 @@
+df <- all_txts_c_dtm
+
+dont_care <- 
+  c('university', 'abstracts', 'meeting', 
+    'annual', 'paper', 'chair', 'across', 
+    'presented', 'session', 'college', 'area',
+    'discussed', 'this', 'that', 'there', 'which',
+    'their', 'poster', 'using', 'through', 'into',
+    'some', 'from', 'been', 'discuss', 'while',
+    'over')
+
+# from tidytext
+df_tbl <- 
+  convert(df, to = "data.frame") %>% 
+  pivot_longer(-document, 
+               names_to = "word",
+               values_to = "n") %>% 
+  mutate(year = parse_number(document)) %>% 
+  filter(word %in% names(data_int_syllables)) %>% 
+  filter(str_length(word) > 3) %>% 
+  filter(!word %in% dont_care)
+
+df_tbl_yearly_totals <- 
+  df_tbl %>% 
+  group_by(year) %>% 
+  summarise(total_words = sum(n, na.rm = TRUE))
+  
+# compute proportion of all words per year
+df_tbl_prop <- 
+  df_tbl %>% 
+ # left_join(all_txts_c_summary) %>% 
+  left_join(df_tbl_yearly_totals) %>% 
+  mutate(prop = n / total_words ) %>% 
+  filter(prop > 0) %>% 
+  group_by(word) %>% 
+  mutate(sum_the_word = sum(n)) %>% 
+  mutate(keyword_n = str_c(word, " (n = ", sum_the_word, ")"))
+
+# want to find words with the greatest difference in 
+# proportion between two time periods 
+
+df_tbl_prop_two_groups <- 
+  df_tbl_prop %>% 
+  mutate(year_group = ifelse(document < 2004, 
+                             "early", 
+                             "late"))
+
+df_tbl_prop_two_groups_diff <- 
+df_tbl_prop_two_groups %>% 
+  group_by(word, year_group) %>% 
+  summarise(mean_prop = mean(prop, na.rm = TRUE)) %>% 
+  pivot_wider(names_from = year_group,
+              values_from = mean_prop) %>% 
+  drop_na() %>% 
+  mutate(diff = late - early)
+
+df_tbl_prop_two_groups_diff_interesting <- 
+df_tbl_prop_two_groups_diff %>% 
+  filter(diff >= 0.00001 | diff <= -0.00001)
+
[…]

### 2021-07-19 code/005-exploring-basic-word-stats-and-plots.R

@@ -1,74 +1,0 @@
-
-
-# Words per year
-
-library(tidyverse)
-library("quanteda")
-
-# get all txt files of abstracts
-abstract_txt_files <- 
-  fs::dir_ls(here::here("data/derived-data/"), 
-             recurse = TRUE, 
-             regexp = '\\.txt$')
-
-all_txts <- tolower(readtext::readtext(abstract_txt_files))
-names(all_txts) <- map_chr(names(all_txts), ~str_match_all(.x, "\\d{4}")[[1]][1,1])
-
-# count all words for each year
-all_txts_c <- corpus(all_txts)
-all_txts_c_summary <- 
-  summary(all_txts_c) %>%   
-  mutate(year = parse_number(Text))
-  
-# make a dfm
-all_txts_c_dtm <- 
-  dfm(all_txts_c,
-      remove_punct = TRUE)
-
-# Explore key words over time
-
-#----- change these words
-keywords <-
-  c("agency",
-    "phenomena",
-    "action",
-    "affect",
-    "inference",
-    "mechanism")
-#----
-
-dfm_keywords <- 
-  dfm_select(all_txts_c_dtm, 
-             pattern = keywords, 
-             selection = "keep")
-
-dfm_keywords_tbl <- 
-  convert(dfm_keywords, to = "data.frame") %>% 
-  pivot_longer(-document, 
-               names_to = "keyword",
-               values_to = "n") %>% 
-  mutate(year = parse_number(document))
-
-# compute proportion of all words per year
-dfm_keywords_tbl_prop <- 
-dfm_keywords_tbl %>% 
-  left_join(all_txts_c_summary) %>% 
-  mutate(prop = n / Tokens )
-
-# plot of keywords as a proportion of all words per year
-ggplot(data = dfm_keywords_tbl_prop, 
-       aes(x = year , 
-           y = prop)) +
-  geom_col() +
-  facet_wrap( ~ keyword, 
-              ncol = 1, 
-             scales = "free_y") +
-  scale_x_continuous(labels = c(seq(1960, 2020, 2)), 
-                     breaks = seq(1960, 2020, 2),
-                     name = "Year") +
-  scale_y_continuous(name = "Proportion of all words per year",
-                     labels = scales::comma) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, 
-                                   vjust = 0.5)) 
-

### 2021-07-19 code/006-temp-ref-scripts.R

@@ -79,7 +79,7 @@ temporal_referencing_plot_output <-
                                time_specific_token = insert_time_specific_token_data$time_specific_token,
                                target_feature,
                                year_interval,
-                               perplexity = 100)
+                               perplexity = 50)
 #-----------------------------------------------------------
 
 # Now we can save the plot:
@@ -93,6 +93,9 @@ ggsave(here::here(str_glue("figures/tr-plot-target-word-is-{target_feature}-{yea
 
 target_words <- 
   c(
+    "statement",
+    "framework",
+    "argument",
  "theory"
 # ,"model"
 # ,"mechanisms"

### 2021-07-19 code/other/999-exploring-basic-word-stats-and-plots.R

@@ -1,0 +1,180 @@
+
+
+# Words per year
+
+library(tidyverse)
+library("quanteda")
+
+# get all txt files of abstracts
+abstract_txt_files <- 
+  fs::dir_ls(here::here("data/derived-data/"), 
+             recurse = TRUE, 
+             regexp = '\\.txt$')
+
+all_txts <- tolower(readtext::readtext(abstract_txt_files))
+names(all_txts) <- map_chr(names(all_txts), ~str_match_all(.x, "\\d{4}")[[1]][1,1])
+
+# count all words for each year
+all_txts_c <- corpus(all_txts)
+all_txts_c_summary <- 
+  summary(all_txts_c) %>%   
+  mutate(year = parse_number(Text))
+  
+# make a dfm
+all_txts_c_dtm <- 
+  dfm(all_txts_c,
+      remove_punct = TRUE)
+
+# Explore key words over time
+
+#----- change these words
+keywords <-
+  c("agency",
+    "phenomena",
+    "action",
+    "affect",
+    "inference",
+    "mechanism")
+#----
+
+dfm_keywords <- 
+  dfm_select(all_txts_c_dtm, 
+             pattern = keywords, 
+             selection = "keep")
+
+dfm_keywords_tbl <- 
+  convert(dfm_keywords, to = "data.frame") %>% 
+  pivot_longer(-document, 
+               names_to = "keyword",
+               values_to = "n") %>% 
+  mutate(year = parse_number(document))
+
+# compute proportion of all words per year
+dfm_keywords_tbl_prop <- 
+dfm_keywords_tbl %>% 
+  left_join(all_txts_c_summary) %>% 
+  mutate(prop = n / Tokens )
+
+# plot of keywords as a proportion of all words per year
+ggplot(data = dfm_keywords_tbl_prop, 
+       aes(x = year , 
+           y = prop)) +
+  geom_col() +
+  facet_wrap( ~ keyword, 
+              ncol = 1, 
+             scales = "free_y") +
+  scale_x_continuous(labels = c(seq(1960, 2020, 2)), 
+                     breaks = seq(1960, 2020, 2),
+                     name = "Year") +
+  scale_y_continuous(name = "Proportion of all words per year",
+                     labels = scales::comma) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, 
+                                   vjust = 0.5)) 
+
+
+##-------------------------------------------------------------------------
+
[…]
